---
title: "2023-24-PSYC122-w17-workbook-answers"
author: Rob Davies
date: "2024-02-26"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction


In Week 17, we aim to *further* develop skills in visualizing and testing the 
associations between variables in psychological data.

We do this to learn how to answer research questions like:

1. What person attributes predict success in understanding?
2. Can people accurately evaluate whether they correctly understand written 
health information?

These kinds of research questions can be answered using methods like the **linear model**.

When we do these analyses, we need to think about how we report the results:

- we usually need to report information about the kind of model we specify;
- and we will need to report the nature of the association estimated in our model;
- we usually need to decide, is the association significant?
- does the association reflect a positive or negative relationship between outcome 
and predictor?
- and is the association we see in our sample data relatively strong or weak?

We will consolidate and extend learning on data visualization:

- focusing on how we edit ggplot() code to produce professional looking plots.

## Naming things

I will format dataset names like this: 

- `study-two-general-participants.csv`

I will also format variable (data column) names like this: `variable` 

I will also format value or other data object (e.g. cell value) names like this: `studyone`

I will format functions and library names like this: e.g. function `ggplot()` or e.g. library `{tidyverse}`.

## The data we will be using

In this activity, we use data from *a second* 2020 study of the response
of adults from a UK national sample to written health information:

- `study-two-general-participants.csv`


# Answers


## Step 1: Set-up


To begin, we set up our environment in R.


### Task 1 -- Run code to empty the R environment

```{r}
rm(list=ls())                            
```


### Task 2 -- Run code to load relevant libraries

```{r}
library("tidyverse")
```


## Step 2: Load the data


### Task 3 -- Read in the data file we will be using

The data file is called:

- `study-two-general-participants.csv`

Use the `read_csv()` function to read the data file into R:

```{r}
study.two.gen <- read_csv("study-two-general-participants.csv")
```

When you read the data file in, give the data object you create a distinct name 
e.g. `study.two.gen`.


### Task 4 -- Inspect the data file

Use the `summary()` or `head()` functions to take a look.

```{r}
head(study.two.gen)
summary(study.two.gen)
```

Even though you have done this before, you will want to do it 
again, here, and pay particular attention to what you see, for the numeric variables, 
in the information about minimum (Min.) and maximum (Max.) values.


## Step 3: Use histograms to examine the distributions of variables


### Revise: practice to strengthen skills


### Task 5 -- Draw histograms to examine the distributions of variables
### hint: Task 5 

Use `ggplot()` with `geom_histogram()`.

When we create a plot, we take things step-by-step.

Here's an example you have seen before: run the lines of code and see the result in the `Plots` window in `R-Studio`.

```{r}
ggplot(data = study.two.gen, aes(x = mean.acc)) + 
  geom_histogram()
```

These are the steps, set out one at a time:

1. `ggplot(...)` you tell R you want to make a plot using the `ggplot()` function
2. `ggplot(data = study.two.gen ...)` you tell R you want to make a plot with the
`study.two.gen` data
3. `ggplot(..., aes(x = mean.acc))` you tell R that you want to make a plot with 
the variable `mean.acc` -- here, you specify the aesthetic mapping, `x = mean.acc`
4. `ggplot(...) + geom_histogram()` you tell R you want to plot values of `mean.acc` 
as a histogram

Notice that the code works the same whether we have the different bits of code on the same
line or in a series of lines.


### Revise: make sure you are confident about doing these things


### Task 6 -- Practice editing the appearance of a histogram plot step-by-step

Start by constructing a basic histogram.

#### hint: Task 6 -- Choose whichever numeric variable from the `study.two.gen` dataset you please

#### hint: Task 6 -- Use the line-by-line format to break the plot code into steps

It will make it easier to read, and it will make it easier to add edits e.g.

```{r}
ggplot(data = study.two.gen, aes(x = SHIPLEY)) + 
  geom_histogram()
```

We are going to revise editing:

1. The appearance of the bars using `binwidth`;
2. The colour of the background using `theme_bw()`;
3. The appearance of the labels using `labs().

Then we are going to try some new moves:

4. Setting the x-axis limits to reflect the full range of possible scores 
on the x-axis variable;
5. Adding annotation -- here, a vertical line -- indicating the sample average 
for a variable.

#### Questions: Task 6

- Q.1. Edit the appearance of the bars by specifying a `binwidth` value.

```{r}
ggplot(data = study.two.gen, aes(x = SHIPLEY)) + 
  geom_histogram(binwidth = 2)
```

- Q.2. Then add an edit to the appearance of the background using `theme_bw()`.

```{r}
ggplot(data = study.two.gen, aes(x = SHIPLEY)) + 
  geom_histogram(binwidth = 2) +
  theme_bw()
```

- Q.3. Then add an edit to the appearance of the labels using `labs()`.

```{r}
ggplot(data = study.two.gen, aes(x = SHIPLEY)) + 
  geom_histogram(binwidth = 2) +
  theme_bw() +
  labs(x = "SHIPLEY", y = "frequency count")
```


### Introduce: make some new moves


- Q.4. Now add an edit by setting the x-axis limits using `x.lim()`.

```{r}
ggplot(data = study.two.gen, aes(x = SHIPLEY)) + 
  geom_histogram(binwidth = 2) +
  theme_bw() +
  labs(x = "Vocabulary (SHIPLEY)", y = "frequency count") +
  xlim(0,40)
```

- Q.5. Then add an edit to draw a vertical line to show the mean value of 
the variable you are plotting.

```{r}
ggplot(data = study.two.gen, aes(x = SHIPLEY)) + 
  geom_histogram(binwidth = 2) +
  theme_bw() +
  labs(x = "Vocabulary (SHIPLEY)", y = "frequency count") +
  xlim(0,40) +
  geom_vline(xintercept = mean(study.two.gen$SHIPLEY), colour = "red", size = 1.5)
```

- Q.6. Can you find information on how to define the limits on the x-axis 
and on the y-axis?
- hint: Q.6. You can see the information in this week's how-to but try a 
search online for "ggplot reference xlim".

- A.6. See ggplot reference information on setting limits here:

<https://ggplot2.tidyverse.org/reference/lims.html>

- Q.7. Can you find information on how to a reference line?
- hint: Q.7. You can see the information in this week's how-to but try a 
search online for "ggplot reference vline".

- A.7. See ggplot reference information on setting limits here:

<https://ggplot2.tidyverse.org/reference/geom_abline.html>


## Step 4: Now draw scatterplots to examine associations between variables 


### Consolidation: should be no surprises here


### Task 7 -- Create a scatterplot to examine the association between some variables 

Between the outcome `mean.acc` and each of three numeric potential predictor variables 
`SHIPLEY`, `HLVA` and `AGE`.

#### hint: Task 7 -- We are working with `geom_point()` and you need x and y aesthetic mappings.
#### hint: Task 7 -- The outcome variable `mean.acc` has to be mapped to the y-axis using `...y = ...`

```{r}
ggplot(data = study.two.gen, aes(x = SHIPLEY, y = mean.acc)) +
  geom_point()
```

```{r}
ggplot(data = study.two.gen, aes(x = HLVA, y = mean.acc)) +
  geom_point()
```

```{r}
ggplot(data = study.two.gen, aes(x = AGE, y = mean.acc)) +
  geom_point()
```


### Revise: make sure you are confident about doing these things


### Task 8 -- Edit the appearance of *each* plot step-by-step

#### hint: Task 8 -- You may want to use the same plot appearance choices for all plots 

Because a consistent appearance is generally neater and easier for 
your audience to process.

#### hint: Task 8 -- You can find links to reference information on options in the how-to guide

Use the information to make the plots pleasing in appearance to you.

#### hint: Task 8 -- Do not be afraid to copy then paste code you re-use.

But be careful that things like axis values are sensible for each variable.

#### Questions: Task 8

- Q.8. -- First, edit the appearance of the points using `alpha`, `size`, `shape`, and `colour`:

```{r}
ggplot(data = study.two.gen, aes(x = SHIPLEY, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')
```

```{r}
ggplot(data = study.two.gen, aes(x = HLVA, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')
```

```{r}
ggplot(data = study.two.gen, aes(x = AGE, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')
```

- Q.9. -- Then edit the colour of the background using `theme_bw()`:

```{r}
ggplot(data = study.two.gen, aes(x = SHIPLEY, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')   +
  theme_bw()
```

```{r}
ggplot(data = study.two.gen, aes(x = HLVA, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')   +
  theme_bw()
```

```{r}
ggplot(data = study.two.gen, aes(x = AGE, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')   +
  theme_bw()
```

- Q.10. -- Then edit the appearance of the labels using `labs()`:

```{r}
ggplot(data = study.two.gen, aes(x = SHIPLEY, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')   +
  theme_bw() +
  labs(x = "SHIPLEY", y = "mean accuracy")
```

```{r}
ggplot(data = study.two.gen, aes(x = HLVA, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')   +
  theme_bw() +
  labs(x = "HLVA", y = "mean accuracy")
```

```{r}
ggplot(data = study.two.gen, aes(x = AGE, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')   +
  theme_bw() +
  labs(x = "Age (Years)", y = "mean accuracy")
```


### Introduce: make some new moves


- Q.11. -- Then set the x-axis and y-axis limits to the minimum-maximum ranges 
of the variables you are plotting
- hint: Q.11. -- For these plots the y-axis limits will be the same because 
the outcome stays the same
- hint: Q.11. -- But the x-axis limits will be different for each different 
predictor variable
- hint: Q.11. -- The minimum value will always be 0

```{r}
ggplot(data = study.two.gen, aes(x = SHIPLEY, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')   +
  theme_bw() +
  labs(x = "SHIPLEY", y = "mean accuracy") +
  xlim(0, 40) + ylim(0, 1)
```

```{r}
ggplot(data = study.two.gen, aes(x = HLVA, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')   +
  theme_bw() +
  labs(x = "HLVA", y = "mean accuracy") +
  xlim(0, 16) + ylim(0, 1)
```

```{r}
ggplot(data = study.two.gen, aes(x = AGE, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')   +
  theme_bw() +
  labs(x = "Age (Years)", y = "mean accuracy") +
  xlim(0, 80) + ylim(0, 1)
```


## Step 5: Use correlation to to answer the research questions 


### Revise: make sure you are confident about doing these things


One of our research questions is:

1. What person attributes predict success in understanding?


### Task 9 -- Examine the correlations between the outcome variable and predictor variables

#### hint: Task 9 -- We use `cor.test()`
#### hint: Task 9 -- You can look at the `how-to` guide for more advice

You need to run three separate correlations: 

1. between mean accuracy and SHIPLEY
2. between mean accuracy and HLVA and 
3. between mean accuracy and AGE

```{r}
cor.test(study.two.gen$SHIPLEY, study.two.gen$mean.acc, method = "pearson",  alternative = "two.sided")
```

```{r}
cor.test(study.two.gen$HLVA, study.two.gen$mean.acc, method = "pearson",  alternative = "two.sided")
```

```{r}
cor.test(study.two.gen$AGE, study.two.gen$mean.acc, method = "pearson",  alternative = "two.sided")
```

Now use the results from the correlations to answer the following questions.

- Q.12. -- What is `r`, the coefficient for the correlation between `mean.acc` and `SHIPLEY`?

- A.12. -- r = 0.4650537

- Q.13. -- Is the correlation between `mean.acc` and `HLVA`  significant?

- A.13. -- r is significant, p < .05

- Q.14. -- What are the values for t and p for the significance test for the correlation between 
`mean.acc` and `AGE`?

- A.14. -- t = 0.30121, p = 0.7636

- Q.15. -- For which pair of outcome-predictor variables is the correlation the largest?

- A.15. -- The correlation is the largest between `mean.acc` and `HLVA`.

- Q.16. -- What is the sign or direction of each of the correlations?

- A.16. -- All the correlations are positive.


## Step 6: Use a linear model to to answer the research questions 


### Introduce: Make some new moves


One of our research questions is:

1. What person attributes predict success in understanding?

### Task 10 -- Examine the relation between outcome mean accuracy (`mean.acc`) 

And each of the predictors: `SHIPLEY`, `HLVA` and `AGE`

#### hint: Task 10 -- Use `lm()`
#### hint: Task 10 -- Run three separate `lm()` analyses

All with `mean.acc` as the outcome but each with one predictor variable

#### hint: Task 10 -- See the `how-to` guide for example code

```{r}
model <- lm(mean.acc ~ SHIPLEY, data = study.two.gen)
summary(model)
```

```{r}
model <- lm(mean.acc ~ HLVA, data = study.two.gen)
summary(model)
```

```{r}
model <- lm(mean.acc ~ AGE, data = study.two.gen)
summary(model)
```

Notice that we do the linear model in the steps:

1. `model <- lm(...)` fit the model using `lm(...)`, and give the model a name,
here, we call it `model`;
2. `...lm(mean.acc ~ SHIPLEY...)` tell R you want a model of 
the outcome `mean.acc` predicted `~` by the predictor `SHIPLEY`
3. `...data = study.two)` tell R that the variables you name in the formula live in the 
`study.two` dataset.
4. `summary(model)` ask R for a summary of the model you called `model`.

Notice that R has a general formula syntax: 

`outcome ~ predictor *or* y ~ x`

- and uses the same format across a number of different functions;
- each time, the left of the tilde symbol `~` is some output or outcome;
- and the right of the tilde `~` is some input or predictor or set of predictors.

#### Questions: Task 10
  
If you look at the model summary you can answer the following questions.

- Q.17. What is the estimate for the coefficient of the effect of the predictor 
`HLVA` on `mean.acc`?

- A.17. 0.026207 

- Q.18. Is the effect significant?

- A.18. It is significant, p < .05

- Q.19. What are the values for t and p for the significance test for the coefficient?

- A.19. t = 7.529, p = 2.87e-12

- Q.20. How would you describe in words the shape or direction of the association 
between `HLVA` and `mean.acc`?

- A.20. The slope coefficient -- and a scatterplot (draw it) -- suggest that 
as HLVA scores increase so also do mean accuracy scores.

- Q.21. How how would you describe the relations apparent between the predictor and 
outcome in all three models?

- A.21. It is possible to see, given coefficient estimates, that the association 
between predictor and outcome is positive for each model: mean accuracy appears to increase 
for increasing values of SHIPLEY vocabulary, HLVA health literacy, and age.


## Step 7: Use a linear model to generate predictions 


### Introduce: make some new moves


### Task 11 -- We can use the model we have just fitted to plot the model predictions

#### hint: Task 11 -- We are going to draw a scatterplot and add a line 

The line will show the model predictions, given the model intercept and effect 
coefficient estimates.

First fit a model and get a summary: model the relationship between `mean.acc` and `HLVA`

```{r}
model <- lm(mean.acc ~ HLVA, data = study.two.gen)
summary(model)
```

- Q.22. What is the coefficient estimate for the intercept?

- A.22. 0.522016

- Q.23. What is the coefficient estimate for the slope of `HLVA` (see earlier)?

- A.23. 0.026207

Second, use the `geom_point()` to draw a scatterplot and `geom_abline()` function 
to draw the prediction line representing the association between this outcome and predictor

```{r}
ggplot(data = study.two.gen, aes(x = HLVA, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = 'square')   +
  geom_abline(intercept = 0.522016, slope = 0.026207, colour = "red", size = 1.5) +
  theme_bw() +
  labs(x = "HLVA", y = "mean accuracy") +
  xlim(0, 15) + ylim(0, 1)
```

You can see reference information here:  

<https://ggplot2.tidyverse.org/reference/geom_abline.html>

