---
title: "2023-24-PSYC122-w17-how-to"
author: Rob Davies
date: "2024-02-26"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction


In Week 17, we aim to *further* develop skills in visualizing and testing the 
associations between variables in psychological data.

We do this to learn how to answer research questions like:

1. What person attributes predict success in understanding?
2. Can people accurately evaluate whether they correctly understand written 
health information?

These kinds of research questions can be answered using methods like the **linear model**.

When we do these analyses, we need to think about how we report the results:

- we usually need to report information about the kind of model we specify;
- and we will need to report the nature of the association estimated in our model;
- we usually need to decide, is the association significant?
- does the association reflect a positive or negative relationship between outcome 
and predictor?
- and is the association we see in our sample data relatively strong or weak?

We will consolidate and extend learning on data visualization:

- focusing on how we edit ggplot() code to produce professional looking plots.

## Naming things

I will format dataset names like this: 

- `study-one-general-participants.csv`

I will also format variable (data column) names like this: `variable` 

I will also format value or other data object (e.g. cell value) names like this: `studyone`

I will format functions and library names like this: e.g. function `ggplot()` or e.g. library `{tidyverse}`.

## The data we will be using

In this activity, we use data from a 2020 study of the response
of adults from a UK national sample to written health information:

- `study-one-general-participants.csv`


## Step 1: Set-up


To begin, we set up our environment in R.


### Task 1 -- Run code to empty the R environment

```{r}
rm(list=ls())                            
```


### Task 2 -- Run code to load relevant libraries

```{r}
library("tidyverse")
```


## Step 2: Load the data


### Task 3 -- Read in the data file we will be using

The data file is called:

- `study-one-general-participants.csv`

Use the `read_csv()` function to read the data file into R:

```{r}
study.one.gen <- read_csv("study-one-general-participants.csv")  
```

When you read the data file in, give the data object you create a distinct name 
e.g. `study.one.gen`.


### Task 4 -- Inspect the data file

Use the `summary()` or `head()` functions to take a look.

```{r}
head(study.one.gen)
summary(study.one.gen)
```

Notice that:

- `head()` will give you the top few rows of any dataset you have read into R.
- `summary()` will give you either descriptive statistics for variable columns
classified as numeric or will tell you that columns in the dataset are not numeric.

Pay attention to what you see, for the numeric variables, in the information about 
minimum (Min.) and maximum (Max.) values. You will use this informatio later when
you engage in data visualization.


## Step 3: Use histograms to examine the distributions of variables


### Revise: practice to strengthen skills


### Task 5 -- Draw histograms to examine the distributions of variables
### hint: Task 5 

Use `ggplot()` with `geom_histogram()`.

When we create a plot, we take things step-by-step.

Here's an example you have seen before: run the lines of code and see the result in the
`Plots` window in `R-Studio`.

```{r}
ggplot(data = study.one.gen, aes(x = mean.acc)) + 
  geom_histogram()
```

These are the steps, set out one at a time:

1. `ggplot(...)` you tell R you want to make a plot using the `ggplot()` function
2. `ggplot(data = study.one.gen ...)` you tell R you want to make a plot with the
`study.two.gen` data
3. `ggplot(..., aes(x = mean.acc))` you tell R that you want to make a plot with 
the variable `mean.acc` -- here, you specify the aesthetic mapping, `x = mean.acc`
4. `ggplot(...) + geom_histogram()` you tell R you want to plot values of `mean.acc` 
as a histogram

Notice that the code works the same whether we have the different bits of code on the same
line or in a series of lines.


### Revise: make sure you are confident about doing these things


### Task 6 -- Practice editing the appearance of a histogram plot step-by-step

We are going to revise editing:

1. The appearance of the bars using `binwidth`;
2. The colour of the background using `theme_bw()`;
3. The appearance of the labels using `labs()`.

We proceed to revise:

#### Task 6.1 -- the appearance of the bars using `binwidth`:

```{r}
ggplot(data = study.one.gen, aes(x = SHIPLEY)) + 
  geom_histogram(binwidth = 2)
```

#### Task 6.2 -- then the colour of the background using `theme_bw()`:

```{r}
ggplot(data = study.one.gen, aes(x = SHIPLEY)) + 
  geom_histogram(binwidth = 2) +
  theme_bw()
```

#### Task 6.3 -- then the appearance of the labels using `labs()`:

```{r}
ggplot(data = study.one.gen, aes(x = SHIPLEY)) + 
  geom_histogram(binwidth = 2) +
  theme_bw() +
  labs(x = "mean accuracy", y = "frequency count")
```


### Introduce: make some new moves


#### Task 6.4. Now add an edit by setting the x-axis limits using `x.lim()`:

```{r}
ggplot(data = study.one.gen, aes(x = SHIPLEY)) + 
  geom_histogram(binwidth = 2) +
  theme_bw() +
  labs(x = "Vocabulary (SHIPLEY)", y = "frequency count") +
  xlim(0,40)
```

#### Task 6.5. Now add a vertical line to show the mean value:

```{r}
ggplot(data = study.one.gen, aes(x = SHIPLEY)) + 
  geom_histogram(binwidth = 2) +
  theme_bw() +
  labs(x = "Vocabulary (SHIPLEY)", y = "frequency count") +
  xlim(0,40) +
  geom_vline(xintercept = mean(study.one.gen$SHIPLEY), colour = "red", size = 1.5)
```

#### Why are we learning how to do this?

It is often useful to define the limits on the x-axis and on the y-axis:

- Because we want to show the audience where the sample values are distributed 
compared to where they *could* be distributed, given the measure.
- This makes sense e.g. if you want to see the relative age distribution of a sample
compared to the population.
- Because we want plots, shown side by side to be directly comparable.
- And because we want to give the audience a more accurate picture of the data.

You can see ggplot reference information on setting limits here:

<https://ggplot2.tidyverse.org/reference/lims.html>

Notice that we used the `geom_vline()` in:

```{}
geom_vline(xintercept = mean(study.one.gen$SHIPLEY), colour = "red", size = 1.5)
```

- to plot a vertical line at the location on the x-axis we define.

We do this in the steps:

1. `geom_vline(...)` draw a vertical line;
2. `...xintercept...` draw the line so it hits the x-axis (intercepts the x-axis);
3. `...xintercept = mean(study.one.gen$SHIPLEY...` defined by the mean of the variable
`mean(study.one.gen$SHIPLEY)`;
4. `...colour = "red", size = 1.5...` make the line red and make the line one and a half times thicker than the default line thickness.

See ggplot() reference information for this new geom is here:

<https://ggplot2.tidyverse.org/reference/geom_abline.html>


## Step 4: Now draw scatterplots to examine associations between variables 


### Consolidation: should be no surprises here


### Task 7 -- Create a scatterplot to examine the association between some variables 

#### hint: Task 7 -- We are working with `geom_point()` and you need x and y aesthetic mappings.
#### hint: Task 7 -- The outcome variable `mean.acc` has to be mapped to the y-axis using `...y = ...`

Run a chunk of code to make the plot.

```{r}
ggplot(data = study.one.gen, aes(x = SHIPLEY, y = mean.acc)) +
  geom_point()
```

This plot shows:

- the possible association between x-axis variable `SHIPLEY` and y-axis 
variable `mean.acc`.

The plot code moves through the following steps:  

1. `ggplot(...)` make a plot;
2. `ggplot(data = study.one.gen, ...)` with the `study.one.gen` dataset;
3. `ggplot(...aes(x = SHIPLEY, y = mean.acc))` using two aesthetic mappings

- `x = SHIPLEY` map `SHIPLEY` values to x-axis (horizontal, left to right) 
positions;
- `y = mean.acc` map `mean.acc` values to y-axis (vertical, bottom to top) 
positions;

4. `geom_point()` show the mappings as points.


### Revise: make sure you are confident about doing these things


### Task 8 -- Edit the appearance of *each* plot step-by-step

We are going to revise making edits to:

1. the appearance of the points using `alpha`, `size`, `shape`, and `colour`;
2. the colour of the background using `theme_bw()`;
3. the appearance of the labels using `labs()`.

We are then going to try out some new moves:

4. Set the x-axis and y-axis limits to the potential minimum-maximum ranges of 
the variables we plot.

Here, "potential" references the fact that e.g. `SHIPLEY` scores can
start at 0 (a person gets 0 responses correct) and has a maximum of 40
(a person could get all responses correct, for up to 40 items).

#### Task 8.1. Edit the appearance of the points using `alpha`, `size`, `shape`, and `colour`:

```{r}
ggplot(data = study.one.gen, aes(x = SHIPLEY, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = "square") 
```

#### Task 8.2. Edit the colour of the background using `theme_bw()`:

```{r}
ggplot(data = study.one.gen, aes(x = SHIPLEY, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = "square")   +
  theme_bw()
```

#### Task 8.3. Edit the appearance of the labels using `labs()`:

```{r}
ggplot(data = study.one.gen, aes(x = SHIPLEY, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = "square")   +
  theme_bw() +
  labs(x = "SHIPLEY", y = "mean accuracy")
```


### Introduce: make some new moves


#### Task 8.4. Set the axis limits:

```{r}
ggplot(data = study.one.gen, aes(x = SHIPLEY, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = "square")   +
  theme_bw() +
  labs(x = "SHIPLEY", y = "mean accuracy") +
  xlim(0, 40) + ylim(0, 1)
```

Notice that I set the x-axis limits to the minimum (0) and maximum (40) possible values
for the `SHIPLEY` variable.
- I set the y-axis limits to the minimum (0) and maximum (1) possible values
for the mean accuracy variable `mean.acc`.

#### Why are we learning how to do this?

It is generally a good idea to show the minimum value (the origin) for each variable.
Not doing this i.e. showing a more narrow slice of the sample range is an easy way to 
*exaggerate* the strength of associations or to *imply incorrectly* the breadth in variation.

You can change the transparency (alpha), size, colour and shape of important parts of a plot.

- Here, we are changing the appearance of the points.
- But you can also change the transparency (alpha), size, colour and shape of reference lines
added to a plot.

The ggplot `geom_point() `reference information is here: 

<https://ggplot2.tidyverse.org/reference/geom_point.html>

- where you can see some examples of the edits we have done.

Some useful information about shape options is here:

<http://www.cookbook-r.com/Graphs/Shapes_and_line_types/>

Some useful information about colourons is here:

<http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/>


### Now experiment!


## Step 5: Use a linear model to to answer the research questions


### Introduce: Make some new moves


One of our research questions is:

1. What person attributes predict success in understanding?

### Task 9 -- Examine the relation between outcome mean accuracy (`mean.acc`) and the predictor variable: vocabulary (`SHIPLEY`)

#### hint: Task 9 -- We use `lm()`

Fit the model and get a summary of the model statistics.

```{r}
model <- lm(mean.acc ~ SHIPLEY, data = study.one.gen)
summary(model)
```

Notice that we do the linear model in the steps:

1. `model <- lm(...)` fit the model using `lm(...)`, give the model a name 
-- here, we call it "model";
2. `...lm(mean.acc ~ SHIPLEY...)` tell R you want a model of the outcome 
`mean.acc` predicted (`~`) by the predictor `SHIPLEY`;
3. `...data = study.one.gen)` tell R that the variables you name in the 
formula live in the `study.one.gen` dataset;
4. `summary(model)` ask R for a summary of the model you called "model".

Notice: R has a general formula syntax: `outcome ~ predictor` *or* `y ~ x`

- and uses the same format across a number of different functions;
- each time, the left of the tilde symbol `~` is some output or outcome;
- and the right of the tilde `~` is some input or predictor or set of predictors.

#### Questions: Task 9

If you look at the model summary you can answer the following questions.

- Q.5. What is the estimate for the coefficient of the effect of the predictor, 
`SHIPLEY`?

- A.5. 0.01050 

- Q.6. Is the effect significant?

- A.6. It is significant, p < .05

- Q.7. What are the values for t and p for the significance test for the 
coefficient?

- A.7. t = 4.585, p = 8.85e-06

- Q.8. What do you conclude is the answer to the research question, 
given the linear model results?

- A.8. The model slope estimate suggests that as `SHIPLEY` scores increase 
so `mean.acc` scores increase also.


## Step 6: Use a linear model to generate predictions 


### Introduce: make some new moves


### Task 10 -- We can use the model we have just fitted to plot the model predictions

#### hint: Task 10 -- We are going to draw a scatterplot and add a line 

The line will show the model predictions, given the model intercept and effect coefficient estimates.

First fit a model and get a summary: model the relationship between `mean.acc` and `SHIPLEY`

```{r}
model <- lm(mean.acc ~ SHIPLEY, data = study.one.gen)
summary(model)
```

- Q.9. What is the coefficient estimate for the intercept?

- A.9. 0.44914

- Q.10. What is the coefficient estimate for the slope of `SHIPLEY`?

- A.10. 0.01050

Second, use the `geom_abline()` function to draw the line:

```{r}
ggplot(data = study.one.gen, aes(x = SHIPLEY, y = mean.acc)) +
  geom_point(alpha = 0.5, size = 2, colour = "blue", shape = "square")   +
  geom_abline(intercept = 0.44914, slope = 0.01050, colour = "red", size = 1.5) +
  theme_bw() +
  labs(x = "SHIPLEY", y = "mean accuracy") +
  xlim(0, 40) + ylim(0, 1)
```

You can see that what we do is:

- add the `geom_abline(...)` function;
- and into that function code, we add information about the intercept and the slope
which we take from the model summary.

You can see reference information here:  

<https://ggplot2.tidyverse.org/reference/geom_abline.html>      

Note that we can get the prediction line drawn for us automatically, as:

```{r}
ggplot(data = study.one.gen, aes(x = SHIPLEY, y = mean.acc)) +
  # geom_point(alpha = 0.5, size = 2, colour = "blue", shape = `square`)   +
  geom_smooth(method = "lm", colour = "purple", alpha = .2, size = 2.5, se = FALSE) +
  geom_abline(intercept = 0.44914, slope = 0.01050, colour = "red", size = 1) +
  theme_bw() +
  labs(x = "SHIPLEY", y = "mean accuracy") +
  xlim(0, 40) + ylim(0, 1)  
```

- Here, I turned the points off by commenting them out, adding # to the start of the line.
- And I added `geom_smooth(method = "lm", ...)` to draw a prediction line.
- You can compare the red prediction line I drew using the model estimates
- with the purple line I used geom_smooth() to draw automatically to see that they are identical

This shows you something of what geom_smooth() does.
It is very useful:

<https://ggplot2.tidyverse.org/reference/geom_smooth.html>

