---
title: 1. Week 1 - Being a scientist, doing a science
subtitle: Written by Rob Davies and Margriet Groen
order: 2 # ensure this is one more than the previous script
---

:::{.callout-note title="Overview"}
This week we focus on two strands of learning:

1. Rob introduces you to what it means to do psychological science.
2. Margriet talks you through how to get started in RStudio.

The materials we have prepared to support your learning are arranged in four parts:

- Lectures
- Reading
- Pre-lab activities
- Lab activities

You can navigate to each in the side bar on the right
:::

<!-- Note: I think it is helpful to have a list of the different sections in a week for cross-referencing. I don't think it helps readability it this is visible to the students (hence I've commented them out) as the outline in the sidebar on the right already provides this info. -->
<!-- - @sec-4141-w1-lectures: Lectures: watch these before the lab session -->
<!-- - @sec-4141-w1-lectures-intro-what-module: What the module is about -->
<!-- - @sec-4141-w1-lectures-lecture: The lecture: Being a scientist, doing a science -->
<!-- - @sec-4141-w1-lectures-intro-how-r: How to get started with RStudio -->
<!-- - @sec-4141-w1-reading: Recommended readings -->
<!-- - @sec-4141-w1-prelab-activities: Activities to do before you attend your lab session -->
<!-- - @sec-4141-w1-lab-activities: Activities you do during your lab session -->

## Lectures {#sec-4141-w1-lectures}

### What this module is about: begin here {#sec-4141-w1-lectures-intro-what-module}

**Introduction to the module** (~18 minutes)

Watch this part first. It gives an overview of what the module is about, how the module is organised, how you'll be assessed and how you can ask for help. The video has captions, in case you find that helpful. You can download the slides and a transcript from the links below the video.

```{r}
# Do not remove. It prevents Quarto from thinking it needs to involve Python.
```

```{=html}
<div style="position: relative; width: 100%; height: 0; padding-bottom: 56.25%">
	<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=e32a54eb-f436-4a2e-9beb-b4d200e6920c&autoplay=true&offerviewer=true&showtitle=false&showbrand=false&captions=true&interactivity=all" style="border: 1px solid #464646; position: absolute; top: 0; left: 0; width: 100%; height: 100%; box-sizing: border-box;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="4141_intro"></iframe>
</div>
```

\newline

- [Intro slides](files/week1/4141_intro_slides.pdf)
- [Intro transcript](files/week1/4141_intro_transcript.txt)

### Interactive lecture -- Being a scientist, doing a science: watch this to get started on learning and thinking about what we are going to be doing in Psychology {#sec-4141-w1-lectures-lecture} 

You can watch the lecture for the video using the video playback windows shown next. The lecture is broken into five parts, each about 15 minutes long (some a bit longer).

1. **Big ideas, difficult questions**
2. **The hardest science**
3. **The logic, the challenges**
4. **Science is a thing that is done by people**
5. **Science is a social activity: fads, inequities and biases**

::: callout-tip
During the lecture, we suggest some questions you can think about. Responding to these questions will help you to learn.

It would be great if you cared to enter your questions to me (Rob) about the lecture. You can do that anonymously by clicking on this link and typing in your thoughts:

<https://app.sli.do/event/gChjffSSfLvkAJSr9GNuCA>
:::

This is a quick (~4 minutes) walk through of how you can watch the video lecture, with tips on how to make playback accessible and convenient.

```{=html}
<div style="position: relative; width: 100%; height: 0; padding-bottom: 56.25%">
	<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=0f42ac15-3d1c-4f3f-8c3a-b4d50115be65&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=true&start=0&interactivity=all" style="border: 1px solid #464646; position: absolute; top: 0; left: 0; width: 100%; height: 100%; box-sizing: border-box;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="video-lecture-how-to"></iframe>
</div>
```

\newline

You can download the lecture slides as a .html file you can open in any browser.

The slides end with a references list but don't worry about reading what's in the references list, we give you recommended readings in @sec-4141-w1-reading.

::: callout-tip
The slides presented in the next videos can be downloaded as a web page: `doing-being-science.html.` 

- [Lecture slides](files/week1/doing-being-science.html)

When you download the file, double click on it to open it in your usual web browser for viewing. In some browsers, clicking on the link will just open the slides as a webpage in your browser.
:::

Next, you can see the lecture elements. We recommend that you watch them in sequence.

1. **Big ideas, difficult questions** (~17 minutes). What is science, how do we do it?

\newline

```{=html}
<div style="position: relative; width: 100%; height: 0; padding-bottom: 56.25%">
	<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=d6a2642d-4374-47e4-9b74-b4d500b1d9fb&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=true&start=0&interactivity=all" style="border: 1px solid #464646; position: absolute; top: 0; left: 0; width: 100%; height: 100%; box-sizing: border-box;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="critical-thinking-one"></iframe>
</div>
```

\newline

2. **The hardest science** (~15 minutes). Why learning about people is hard, and why revolutionary new ways of doing psychological science are our response to the challenges.

```{=html}
<div style="position: relative; width: 100%; height: 0; padding-bottom: 56.25%">
	<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=70c5bbb0-e2b4-4126-8d6c-b4d500b9f6fc&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=true&interactivity=all" style="border: 1px solid #464646; position: absolute; top: 0; left: 0; width: 100%; height: 100%; box-sizing: border-box;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="critical-thinking-two"></iframe>
</div>
```

\newline

3. **The logic, the challenges** (~25 minutes). The logic of the scientific method, and the challengesin doing science when it  involves trying to understand people.

```{=html}
<div style="position: relative; width: 100%; height: 0; padding-bottom: 56.25%">
	<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=2ce3cb67-f8c9-4fb2-bd76-b4d500bf7147&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=true&interactivity=all" style="border: 1px solid #464646; position: absolute; top: 0; left: 0; width: 100%; height: 100%; box-sizing: border-box;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="critical-thinking-three"></iframe>
</div> 
```

\newline

4. **Science is a thing that is done by people** (~18 minutes). How is it that groups of people who may not know or like or trust each other can produce this thing: science?

```{=html}
 <div style="position: relative; width: 100%; height: 0; padding-bottom: 56.25%">
	<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=74066340-0d62-4c7d-a48c-b4d500c982d8&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=true&interactivity=all" style="border: 1px solid #464646; position: absolute; top: 0; left: 0; width: 100%; height: 100%; box-sizing: border-box;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="critical-thinking-four"></iframe>
</div>
```

\newline

5. **Science is a social activity: fads, inequities and biases** (15 minutes). What do we study? Who does the studying? Fads, inequities and biases, and taking responsibility.

```{=html}
<div style="position: relative; width: 100%; height: 0; padding-bottom: 56.25%">
	<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=352d6ea9-7a1f-4ef0-a369-b4d500cf03f3&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=true&interactivity=all" style="border: 1px solid #464646; position: absolute; top: 0; left: 0; width: 100%; height: 100%; box-sizing: border-box;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="critical-thinking-five"></iframe>
</div>
```

\newline

### Getting started with your practical work: you need this for the labs {#sec-4141-w1-lectures-intro-how-r}

1. **How to work with R-Studio**(~16 minutes) 

```{=html}
<div style="position: relative; width: 100%; height: 0; padding-bottom: 56.25%">
	<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=991c9fde-4fb9-4f3e-b4a9-b4d500dbfbe2&autoplay=true&offerviewer=true&showtitle=false&showbrand=false&captions=true&start=2&interactivity=all" style="border: 1px solid #464646; position: absolute; top: 0; left: 0; width: 100%; height: 100%; box-sizing: border-box;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="4141_rstudio"></iframe>
</div>
```

\newline

- [RStudio walkthrough transcript](files/week1/4141_rstudio_transcript.txt)

2. **How to create a plot**(~ minutes)

```{html}
```

\newline

- [How to create a plot walkthrough transcript](files/week1/4141_how_to_plot_transcript.txt)

## Reading {#sec-4141-w1-reading}

::: {.callout-tip title="Easy access to readings"}
All reading is accessible via the library. You can go to Moodle -> Library Resources -> Resource List or you can follow [**this link**](https://lancaster.alma.exlibrisgroup.com/leganto/nui/lists/93496942590001221). Readings have one of the following tags:

- **Essential**: We expect you to be familiar with these reading. You will be assessed on this.
- **Recommended**: We highly recommend these readings. They often provide a different perspective, or explain things in a different way.
- **Optional**: Further readings for those interested.

The reading list is divided into multiple sections:

- **Reading associated with statistics**: this contains links to readings useful for the development of statistical concepts and skills. The main textbook  *Fundamental statistics for the behavioral sciences* by Howell (Essential) lives here.
- **Reading associated with psychological research principles and design**: this contains links to readings that provide a good basis for foundational knowledge around research design.
- **Reading associated with Open Science principles and practices**: this contains links to key literature on the Open Science movement.
- **Week by week**: these contain links to key reading for that week that are not listed on any of the more general lists above.
:::

For **week 1**, the reading is:

- McMillan, B. T. M., Baumgartner, H.A., Bergmann, C. et al. (2026). What 5000 babies can tell us about developing minds and how to study them. *Communications Psychology, 4*, 1-5. **Essential**
- Peterson, D. (2016). The baby factory: Difficult research objects, disciplinary standards, and the production of statistical significance. *Socius, 2*, 1-10. **Recommended**
- Wright, D. B., Barnett, J., & Breakwell, G. M. (2020). Psychology as a science: What is science? In G. M. Breakwell, D. B. Wright, & J. Barnett (Eds.), *Research methods in psychology* (5th ed., pp. 119–138). SAGE Publications. **Recommended**


## Pre-lab activities {#sec-4141-w1-prelab-activities}

We often ask you to do a few things in RStudio before you attend your lab session. Not this week though, so you are ready for the in-person lab session. Well done!

## Lab activities {#sec-4141-w1-lab-activities}

### Lab activity 1:

### Lab activity 2: Logging into the R-server


### Lab activity 3: Reading in data and creating a plot

#### Background

We've all been there: you may be trying to focus on a lecture, but a thought on what you would like to have for lunch pops into your mind. Mind wandering describes the common experience of one's thoughts driting away form the task at hand towards self-generated thoughts and feelings (Smallwood & Schooler, 2006). Today, we'll work with some publicly available data from a research project on mind wandering. Anna Schubert and colleagues (Schubert et al., 2024) conducted a study in which they created a 12-item questionnaire that measures how often your mind wanders without meaning to, how often you let it wander on purpose, and how aware you are when it happens.

#### Aims of the activity

We are going to:

1.  Setup the appropriate bits and pieces in RStudio that we need.
2.  Download the data
3.  Read in the data and have a quick look what is in the file
4.  Create a plot

:::{.callout-important title="Watch the 'How to' lecture"}
If you haven't done so already, this is a good time to watch the 'How to' video (here @sec-4141-w1-lectures) on 'How to create a plot'.
:::

#### Step 1: Setup

Before we can do anything, we need to tell R what packages we want to use. You may wonder what a package is. If R is your kitchen with basic equipment, a package is a specialist cookbook that adds new recipes. All the packages that you'll need have already been installed on the R-server, but you have to tell R which cookbooks to get off the shelf for what you want to do right now. We use the function 'library' to do that. It is useful to have a section at the top of your script where you get all the cookbooks you need for this script off the shelf (...or load all the relevant libraries).

```{r}
# Load relevant libraries
library(here) # the 'here' package helps with finding your files
library(tidyverse) # the 'tidyverse' is a set of packages that handles almost everything you need to do with data
```

#### Step 2: Download the data

Now we need to get the data onto our computer and into our working directory on the R-server.

Download the [data_bmw3.csv](PSYC4141/files/week1/data_bmw3.csv) file by clicking on it. Usually it will be saved in your 'Downloads' folder. Now upload the file into the new folder called 'week 1' in RStudio Server you created during the previous lab activity.

#### Step 3: Reading in the data

The next step is to read in a data file. Here we'll use a subset of the data collected by Schubert and colleagues. It contains responses from 215 students on a questionnaire that measures mind wandering. It is a file of the type `.csv` (which means that individual bits of data are separated by comma's), so we'll use the function `read_csv()` to do this. Inside the function, we tell R where to find the file. This is the bit that comes after `here()`: It specifies the filepath. After the comma, we have the file name, including the extension.

```{r}
# Read in the data
data_bmw3 <- read_csv(here("PSYC4141/files/week1/", "data_bmw3.csv"))
```

#### Step 4: Create a plot

The last thing we'll do today is create our first plot. Let's say we'd like to visualise how many people selected each response option for question 1 of the questionnaire. Question 1 was "While listening to a presentation, my thoughts start to trail off unintentionally." Participants were asked to decide to what extent they agreed or disagreed with that statement. They chose one of the following options: "Fully disagree", "Somewhat disagree", "Neutral", "Somewhat agree" or "Fully agree".

We can do that by creating a histogram. The function we use for that (and for a lot of other plots you'll make for research methods and stats) is `ggplot()`. First we need- simple histogram - change the background - add a label for the x-axis - change the colour (fill) - colours (in the Console) - add labels for the answer options

```{r}
#| label: A very simple histogram
ggplot(data = data_bmw3, mapping = aes(x = BMW3_UN_01)) + # specifies data and variable to plot
  geom_histogram()                                       # specifies what type of plot
```

Now, this does the job, but it doesn't quite look like I'd like it to look. Let's change that. I don't particularly like the grey background. I can change that by adding a layer to the plot specifying a different background. There are lots of different 'themes' in `ggplot()` that you can use. Have a look [here](https://ggplot2.tidyverse.org/reference/ggtheme.html).

```{r}
#| label: With a different background
ggplot(data = data_bmw3, mapping = aes(x = BMW3_UN_01)) + # specifies data and variable to plot
  geom_histogram()  +                                     # specifies what type of plot
  theme_bw()                                              # changes the background
```

I would like the bars to be red instead of grey. I can make that happen, but specifying an additional bit of code (or 'argument' in the `ggplot()` part. There are lots of different colours and colour schemes available within `ggplot()`. Have a look at [this overview](https://r-charts.com/colors/) and choose one you like for your plot. There is more to say about which colour schemes to use, but we'll save that for another day.

```{r}
#| label: The bars are now red
ggplot(data = data_bmw3, mapping = aes(x = BMW3_UN_01, fill = "tomato")) + # ...and makes the bars red
  geom_histogram()  +                                                      # specifies what type of plot
  theme_bw()                                                               # changes the background
```

To avoid the legend on the right getting printed (because we don't need it in this case), we can tell R not to include it

```{r}
#| label: Remove the legend
ggplot(data = data_bmw3, mapping = aes(x = BMW3_UN_01, fill = "tomato")) + # ...and makes the bars red
  geom_histogram()  +                                                      # specifies what type of plot
  theme_bw() +                                                             # changes the background
  theme(legend.position = "none")                                          # removes the legend
```

Now it would be good to have a more informative label for what we've plotted on the x-axis. Instead of the variable name `BMW3_UN_01`, we'll add 'Brief Mind Wandering Three-factor Scale - Question 1'.

```{r}
#| label: Add informative x-axis label
ggplot(data = data_bmw3, mapping = aes(x = BMW3_UN_01, fill = "tomato")) + # ...and makes the bars red
  geom_histogram()  +                                                      # specifies what type of plot
  theme_bw() +                                                             # changes the background
  theme(legend.position = "none") +                                        # removes the legend
  labs(x = "Brief Mind Wandering Three-Factor Scale - Question 1")        # add more informative label for the x-axis
```

Finally, instead of the numerical values 0 to 4 that are currently on the x-axis, we'll add the labels that are associated with each of these values.

```{r}
#| label: Add text labels to x-axis
ggplot(data = data_bmw3, mapping = aes(x = BMW3_UN_01, fill = "tomato")) + # ...and makes the bars red
  geom_histogram()  +                                                      # specifies what type of plot
  theme_bw() +                                                             # changes the background
  theme(legend.position = "none") +                                        # removes the legend
  labs(x = "Brief Mind Wandering Three-Factor Scale - Question 1") +       # add more informative label for the x-axis
  scale_x_continuous(                                                      # replace numerical values with text labels
    breaks = 0:4,
    labels = c("Fully disagree", "Somewhat disagree", "Neutral", "Somewhat agree", "Fully agree")
  )
```

That's it! You have created your first plot.

For more practice, go through the same steps for a different question. Remember, you can look at the questionnaire [here](PSYC4141/files/week1/"BMW-3_eng.pdf")