---
title: "122_wk13_labActivity2"
author: "Margriet Groen"
date: "2024-01-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab activity 2 - Statistics anxiety and engagement in module activities

>*Question 1a* If our hypothesis is correct, what type of correlation (if any) should we observe between students' mean anxiety levels and the variable `n_weeks`? **A negative correlation**

# Step 1. Loading the relevant libraries
```{r libraries}
library(broom)
library(car)
library(tidyverse)
```
# Step 2. Read in the data
```{r read in data}
stars <- read_csv("stars2.csv")
engage <- read_csv("psess.csv")

head(stars) # Look at the data frames
head(engage)
```
>*Question 2a* In the `stars` data frame, what do the numbers in the first row across the three columns refer to? **ID = 3, Question = Q01 and Score = 1 shows us that participant 3 reported a score of 1 on question 1.**

# Step 3. Getting the data ready.
```{r mean anxiety scores}
stars_means <- stars %>%
  group_by(ID) %>%
  summarise(mean_anxiety = mean(Score, na.rm = TRUE))
stars_means
```

>*Question 3a* What is the mean anxiety score for participant 3? **1.058824**

```{r join data frames}
joined <- inner_join(stars_means, engage, "ID")
```

# Step 4: Calculating descriptive statistics

```{r descriptives}
descriptives <- joined %>%
  summarise(mean_anx = mean(mean_anxiety, na.rm = TRUE),
            sd_anx = sd(mean_anxiety, na.rm = TRUE),
            mean_weeks = mean(n_weeks, na.rm = TRUE),
            sd_weeks = sd(n_weeks, na.rm = TRUE))
descriptives
```

>*Question 4a* What are the means and standard deviation for anxiety and engagement with the statistics module? **Anxiety *M* = 2.08, *SD* = 0.56; Engagement *M* = 4.54, *SD* = 2.42.**

# Step 5: Visualise the data

```{r scatterplot}
ggplot(joined, aes(x = mean_anxiety, y = n_weeks)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Anxiety", y = "Engagement") +
  theme_bw()
```

>*Question 5a* What does the scatterplot suggest about the relationship between anxiety and engagement? **That there might indeed be a relatively strong negative correlation between the two; students with more anxiety, engage less.**

# Step 6: Build the regression model ----------------------------------

```{r model}
mod <- lm(n_weeks ~ mean_anxiety, data = joined)
mod_summary <- summary(mod)
mod_summary
```

>*Question 6a* What is the estimate of the y-intercept for the model, rounded to three decimal places? **9.057. Explanation: In the summary table, this is the estimate of the intercept.**

>*Question 6b* To three decimal places, if the General Linear Model for this model is `Y=beta0 + beta1X + e`, then the value of `beta1` is ... **-2.173. Explanation: In the summary table, this is the estimate of mean_anxiety, i.e., the slope.**

>*Question 6c* To three decimal places, for each unit increase in anxiety, engagement decreases by ... **2.173. Explanation: In the summary table, this is also the estimate of mean_anxiety, the slope is how much it decreases so you just remove the - sign.**

>*Question 6d* To two decimal places, what is the overall F-value of the model? **11.99. Explanation: In the summary table, the F-ratio is noted as the F-statistic.**

>*Question 6e* Is the overall model significant? **Yes. Explanation: The overall model p-value is .001428 which is less than .05, therefore significant.**

>*Question 6f* What proportion of the variance does the model explain? **25.52%. Explanation: The variance explained is determined by R-squared, you simply multiple it by 100 to get the percent.**

# Step 7: Checking assumptions

```{r assumptions}
crPlots(mod)                  # Plot linear line and line that best fits the data to check the relationship between outcome and predictor is linear

qqPlot(mod$residuals)         # Create qq-plot to check residuals are normally distributed

residualPlot(mod)             # Create residual plot to check residual show homoscedasticity
```

>*Question 7a* Does the relationship appear to be linear? **Yes, the pink line roughly falls across the dashed blue line and looks mostly linear.**

>*Question 7b* Do the residuals show normality? **Yes, in the qq-plot the open circles mostly assemble around the solid blue line, and fall mostly within the range of the dashed blue lines.**

>*Question 7c* Do the residuals show homoscedasticity? **Yes, the residual plot shows that the spread of the residuals is roughly similar for different fitted values.**

# Step 8: Write up

**A simple linear regression was performed with engagement (*M* = 4.54, *SD* = 0.56) as the outcome variable and statistics anxiety (*M* = 2.08, *SD* = 0.56) as the predictor variable. The results of the regression indicated that the model significantly predicted course engagement (*F*(1, 35) = 11.99, *p* \< .001, *R\^2* = 0.25), accounting for 25% of the variance. Anxiety was a significant negative predictor (*beta* = -2.17, *p* \< 0.001): as anxiety increased, course engagement decreased.**