---
title: "wk11"
output: html_document
date: "2025-01-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab activity 2

# Step 1: Load libraries
```{r}
library(broom)
library(tidyverse)
```

# Step 2: Read in the data
```{r}
mh <- read_csv("MillerHadenData.csv")
mh #just to have a quick look at the data
```

# Step 3: Make a scatterplot
```{r}
ggplot (mh, aes (x = TV, y = Home)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() +
  labs(x = "Time spend reading at home", y = "Time spend watching TV at home")
```

# Step 4: Conduct a correlation analysis
```{r}
results <- cor.test(mh$Home, 
                    mh$TV, 
                    method = "pearson", 
                    alternative = "two.sided") %>% 
  tidy()
results


r <- results %>%
  pull(estimate) %>%
  round(2)

p <- results %>%
  pull(p.value) %>%
  round(3)

df <- results %>%
  pull(parameter)

```
# Step 5: Calculate the coefficient of determination (also called R-squared)

```{r}
rsquared <- r*r
rsquaredPercent <- round(rsquared * 100, 0)
```

# Step 6: Write up the results.

A Pearson’s correlation coefficient was used to assess the relationship between time spent watching TV and time spent reading at home. There was a significant negative correlation, r(23) = -.65, p < .001. As time spent watching TV increased, time spent reading at home decreased.

## Lab activity 3: Hazardours alcohol use and impulsivity.

# Step 1: Read in the data
```{r}
data <- read_csv("alcoholUse_Impulsivity.csv")
data
```

# Step 2: Plot the relationship between hazard alcohol use and impulsivity using a scatterplot and a line of best fit
```{r}
ggplot(data, aes(x = hau, y = imp)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() +
  labs(x = "Hazardous Alcohol Use", y = "Impulsivity")
```

# Step 3: Conduct a correlation analysis, using Pearson's r
```{r}
results <- cor.test(data$hau, 
                    data$imp, 
                    method = "pearson", 
                    alternative = "two.sided") %>% 
  tidy()

results
```

Pull out Pearson's r, the degrees of freedom and the p-value for reporting the results

```{r}
r <- results %>%
  pull(estimate) %>%
  round(2)

df <- results %>%
  pull(parameter)

pvalue <- results %>%
  pull(p.value) %>%
  round(3)

rsquared <- r*r
rsquaredPercent <- round(rsquared * 100, 0)
```

## Step 4: Write up the results
A Pearson’s correlation coefficient was used to assess the relationship alcohol use and impulsivity. There was a significant positive correlation, r(18) = .54, p < .014. As alcohol use increased, impulsivity also increased.


