---
title: "How to build a plot"
author: "Dr. Margriet Groen"
format:
  revealjs:
    theme: simple
    slide-number: true
    transition: fade
    incremental: false
    view-distance: 999
---

```{=html}
<style>
.reveal h2 { font-size: 1.3em; }
</style>
```

## Overview

-   What is a Quarto document and how do we use it?
-   How to run code to read in data and build a plot

::: notes
In this video we'll talk about what a Quarto document is and we'll start looking at how you can use it. I'll also walk through how we use code to get R to read in data and build a plot
:::

## Step 2: Setup - load libraries

```{r}
#| label: load_libraries
#| echo: true
#| eval: false
#| code-line-numbers: true
library(here)
library(tidyverse)
```

1.  The `here` package helps with finding your files.
2.  The `tidyverse` is a set of packages that handles almost everything you need to do with data.

Learn more about the `tidyverse` here: [https://tidyverse.org](https://tidyverse.org)

## Step 3: Reading in the data

```{r}
#| label: read_data
#| echo: true
#| eval: false
#| code-line-numbers: true
data_mw <- read_csv(here("PSYC4141/files/week1/", "data_bmw3.csv"))
```

1. `read_csv` is function that reads in data that are stored in a file of the type `.csv`. We need to tell R where the file is and what it is called.

`<-` is an assignment arrow, telling R that we want to put the data in an object (or dataframe) called `data_mw`

## Step 4: Build the plot

::: panel-tabset

### Code

```{r}
#| label: plot-code
#| echo: true
#| eval: false
#| code-line-numbers: true
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01))
```
1. We use `ggplot()` to tell R that we want to create a plot and we tell R the data the use and the variable to plot

### Plot

```{r}
#| label: plot-output
#| echo: false
#| eval: true
#| fig-width: 5
library(here)
library(tidyverse)
data_mw <- read_csv(here("PSYC4141/files/week1/", "data_bmw3.csv"))
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01)) 
```

:::

## Step 4: a simple histogram

::: panel-tabset

### Code

```{r}
#| label: simple_histogram-code
#| echo: true
#| eval: false
#| code-line-numbers: true
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01)) +   
  geom_histogram()
```

2. We tell R what type of plot to use

::: notes
If you just had the code on line 1, you would not see a plot.
You have to tell R what type of plot to use. Here we'll use a 'histogram', so we tell it to use the geometric object 'histogram'. There are many different types of geometric objects. We'll encounter a few during this module. 
Plus adds a layer
:::

### Plot

```{r}
#| label: simple_histogram-output
#| echo: false
#| eval: true
#| fig-width: 5
library(here)
library(tidyverse)
data_mw <- read_csv(here("PSYC4141/files/week1/", "data_bmw3.csv"))
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01)) +
  geom_histogram()
```

:::

## Step 4: Build the plot - different background

::: panel-tabset

### Code

```{r}
#| label: different_background-code
#| echo: true
#| eval: false
#| code-line-numbers: true
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01)) +
  geom_histogram()  +                                     
  theme_bw()                                             
```

3. Changes the background

::: notes
There are many different themes. Here I've gone for a simple black and white theme.
Note: layer
:::

### Plot

```{r}
#| label: different_background-output
#| echo: false
#| eval: true
#| fig-width: 5
library(here)
library(tidyverse)
data_mw <- read_csv(here("PSYC4141/files/week1/", "data_bmw3.csv"))
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01)) +
  geom_histogram() +
  theme_bw()
```
:::

## Step 4: Build the plot - change the colours of the bars

::: panel-tabset

### Code

```{r}
#| label: colour_bars-code
#| echo: true
#| eval: false
#| code-line-numbers: true
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01, fill = "tomato")) + 
  geom_histogram()  + 
  theme_bw()
```

1. Specify an additional bit of code (or `argumment`) inside the `mapping` part of the function `ggplot()`. Instead of using the default grey, I'm telling R to 'fill in' the bars with a tomato red.

### Plot

```{r}
#| label: colour_bars-output
#| echo: false
#| eval: true
#| fig-width: 5
library(here)
library(tidyverse)
data_mw <- read_csv(here("PSYC4141/files/week1/", "data_bmw3.csv"))
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01, fill = "tomato")) +
  geom_histogram() +
  theme_bw()
```
:::

## Step 4: Build the plot - remove the legend

::: panel-tabset

### Code

```{r}
#| label: remove_legend-code
#| echo: true
#| eval: false
#| code-line-numbers: true
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01, fill = "tomato")) + 
  geom_histogram()  + 
  theme_bw() +
  theme(legend.position = "none")
```

4. Removes the legend

### Plot

```{r}
#| label: remove_legend-output
#| echo: false
#| eval: true
#| fig-width: 5
library(here)
library(tidyverse)
data_mw <- read_csv(here("PSYC4141/files/week1/", "data_bmw3.csv"))
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01, fill = "tomato")) +
  geom_histogram() +
  theme_bw() +
  theme(legend.position = "none")
```
:::

## Step 4: Build the plot - change the title of the x-axis

::: panel-tabset

### Code

```{r}
#| label: title_x_axis-code
#| echo: true
#| eval: false
#| code-line-numbers: true
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01, fill = "tomato")) + 
  geom_histogram()  + 
  theme_bw() +
  theme(legend.position = "none") +
  labs(x = "Brief Mind Wandering Three-Factor Scale - Question 1")
```

5.  Adds a more informative title for the x-axis.

### Plot

```{r}
#| label: title_x_axis-output
#| echo: false
#| eval: true
#| fig-width: 5
library(here)
library(tidyverse)
data_mw <- read_csv(here("PSYC4141/files/week1/", "data_bmw3.csv"))
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01, fill = "tomato")) +
  geom_histogram() +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x = "Brief Mind Wandering Three-Factor Scale - Question 1")
```
:::

## Step 4: Build the plot - change the value labels

::: panel-tabset

### Code

```{r}
#| label: add_text_labels_x-axis-code
#| echo: true
#| eval: false
#| code-line-numbers: true
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01, fill = "tomato")) + 
  geom_histogram()  + 
  theme_bw() +
  theme(legend.position = "none") +
  labs(x = "Brief Mind Wandering Three-Factor Scale - Question 1") +
  scale_x_continuous(
    breaks = 0:4,
    labels = c("Fully disagree", "Somewhat disagree", "Neutral", "Somewhat agree", "Fully agree")
  )
```

Lines 6. to 9.  Replaces numerical values with text labels

### Plot

```{r}
#| label: add_text_labels_x_axis-output
#| echo: false
#| eval: true
#| fig-width: 5
library(here)
library(tidyverse)
data_mw <- read_csv(here("PSYC4141/files/week1/", "data_bmw3.csv"))
ggplot(data = data_mw, mapping = aes(x = BMW3_UN_01, fill = "tomato")) +
  geom_histogram() +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x = "Brief Mind Wandering Three-Factor Scale - Question 1") +
  scale_x_continuous(
    breaks = 0:4,
    labels = c("Fully disagree", "Somewhat disagree", "Neutral", "Somewhat agree", "Fully agree")
  )
```
