---
title: "4141-w1-lab_act1"
editor: visual
---

## Background

We've all been there. I may be trying to focus on a lecture, but a thought on what you would like to have for lunch pops into your mind. Mind wandering describes the common experience of one's thoughts driting away form the task at hand towards self-generated thoughts and feelings (Smallwood & Schooler, 2006). Today, we'll work with some publicly available data from a research project on mind wandering. Anna Schubert and colleagues (Schubert et al., 2024) conducted a study in which they created a 12-item questionnaire that measures how often your mind wanders without meaning to, how often you let it wander on purpose, and how aware you are when it happens.

## Aims of the activity

We are going to:

1.  Setup the appropriate bits and pieces in RStudio that we need.
2.  Download the data
3.  Read in the data and have a quick look what is in the file
4.  Create a plot

::: callout-important
## Watch the 'How to' video

If you haven't done so already, this is a good time to watch the 'How to' video (here @sec-4141-w1-lectures) on 'How to create a plot'.
:::

## Step 1: Setup

Before we can do anything, we need to tell R what packages we want to use. You may wonder what a package is. If R is your kitchen with basic equipment, a package is a specialist cookbook that adds new recipes. All the packages that you'll need have already been installed on the R-server, but you have to tell R which cookbooks to get off the shelf for what you want to do right now. We use the function 'library' to do that. It is useful to have a section at the top of your script where you get all the cookbooks you need for this script off the shelf (...or load all the relevant libraries).

```{r}
# Load relevant libraries
library(here) # the 'here' package helps with finding your files
library(tidyverse) # the 'tidyverse' is a set of packages that handles almost everything you need to do with data
```

## Step 2: Download the data

Now we need to get the data onto our computer and into our working directory on the R-server.

Download the [data_bmw3.csv](PSYC4141/files/week1/data_bmw3.csv) file by clicking on it. Usually it will be saved in your 'Downloads' folder. Now upload the file into the new folder called 'week 1' in RStudio Server you created during the previous lab activity.

## Step 3: Reading in the data

The next step is to read in a data file. Here we'll use a subset of the data collected by Schubert and colleagues. It contains responses from 215 students on a questionnaire that measures mind wandering. It is a file of the type `.csv` (which means that individual bits of data are separated by comma's), so we'll use the function `read_csv()` to do this. Inside the function, we tell R where to find the file. This is the bit that comes after `here()`: It specifies the filepath. After the comma, we have the file name, including the extension.

```{r}
# Read in the data
data_bmw3 <- read_csv(here("PSYC4141/files/week1/", "data_bmw3.csv"))
```

## Step 4: Create a plot

The last thing we'll do today is create our first plot. Let's say we'd like to visualise how many people selected each response option for question 1 of the questionnaire. Question 1 was "While listening to a presentation, my thoughts start to trail off unintentionally." Participants were asked to decide to what extent they agreed or disagreed with that statement. They chose one of the following options: "Fully disagree", "Somewhat disagree", "Neutral", "Somewhat agree" or "Fully agree".

We can do that by creating a histogram. The function we use for that (and for a lot of other plots you'll make for research methods and stats) is `ggplot()`. First we need- simple histogram - change the background - add a label for the x-axis - change the colour (fill) - colours (in the Console) - add labels for the answer options

```{r}
#| label: A very simple histogram
ggplot(data = data_bmw3, mapping = aes(x = BMW3_UN_01)) + # specifies data and variable to plot
  geom_histogram()                                       # specifies what type of plot
```

Now, this does the job, but it doesn't quite look like I'd like it to look. Let's change that. I don't particularly like the grey background. I can change that by adding a layer to the plot specifying a different background. There are lots of different 'themes' in `ggplot()` that you can use. Have a look [here](https://ggplot2.tidyverse.org/reference/ggtheme.html).

```{r}
#| label: With a different background
ggplot(data = data_bmw3, mapping = aes(x = BMW3_UN_01)) + # specifies data and variable to plot
  geom_histogram()  +                                     # specifies what type of plot
  theme_bw()                                              # changes the background
```

I would like the bars to be red instead of grey. I can make that happen, but specifying an additional bit of code (or 'argument' in the `ggplot()` part. There are lots of different colours and colour schemes available within `ggplot()`. Have a look at [this overview](https://r-charts.com/colors/) and choose one you like for your plot. There is more to say about which colour schemes to use, but we'll save that for another day.

```{r}
#| label: The bars are now red
ggplot(data = data_bmw3, mapping = aes(x = BMW3_UN_01, fill = "tomato")) + # ...and makes the bars red
  geom_histogram()  +                                                      # specifies what type of plot
  theme_bw()                                                               # changes the background
```

To avoid the legend on the right getting printed (because we don't need it in this case), we can tell R not to include it

```{r}
#| label: Remove the legend
ggplot(data = data_bmw3, mapping = aes(x = BMW3_UN_01, fill = "tomato")) + # ...and makes the bars red
  geom_histogram()  +                                                      # specifies what type of plot
  theme_bw() +                                                             # changes the background
  theme(legend.position = "none")                                          # removes the legend
```

Now it would be good to have a more informative label for what we've plotted on the x-axis. Instead of the variable name `BMW3_UN_01`, we'll add 'Brief Mind Wandering Three-factor Scale - Question 1'.

```{r}
#| label: Add informative x-axis label
ggplot(data = data_bmw3, mapping = aes(x = BMW3_UN_01, fill = "tomato")) + # ...and makes the bars red
  geom_histogram()  +                                                      # specifies what type of plot
  theme_bw() +                                                             # changes the background
  theme(legend.position = "none") +                                        # removes the legend
  labs(x = "Brief Mind Wandering Three-Factor Scale - Question 1")        # add more informative label for the x-axis
```

Finally, instead of the numerical values 0 to 4 that are currently on the x-axis, we'll add the labels that are associated with each of these values.

```{r}
#| label: Add text labels to x-axis
ggplot(data = data_bmw3, mapping = aes(x = BMW3_UN_01, fill = "tomato")) + # ...and makes the bars red
  geom_histogram()  +                                                      # specifies what type of plot
  theme_bw() +                                                             # changes the background
  theme(legend.position = "none") +                                        # removes the legend
  labs(x = "Brief Mind Wandering Three-Factor Scale - Question 1") +       # add more informative label for the x-axis
  scale_x_continuous(                                                      # replace numerical values with text labels
    breaks = 0:4,
    labels = c("Fully disagree", "Somewhat disagree", "Neutral", "Somewhat agree", "Fully agree")
  )
```

That's it! You have created your first plot.

For more practice, go through the same steps for a different question.
