---
title: 8. Week 19 -- Linear models -- interactions
subtitle: Written by Rob Davies
order: 9
---

<!-- ::: {.callout-caution} -->
<!-- This page is under construction for 24/25 and may be subject to change before the teaching week. -->
<!-- ::: -->

::: callout-tip
- Here is a link to the sign-in page for [R-Studio Server](https://psy-rstudio.lancaster.ac.uk/auth-sign-in?appUri=%2F)
:::

```{r}
#| warning: false
#| echo: false
library("ggeffects")
library("tidyverse")
```

```{r}
#| warning: false
#| echo: false
all.data <- read_csv("all.data.csv")  
```

## Week 19: Introduction {#sec-wk19-introduction}

Welcome to your overview of our work together in **PSYC122 Week 19**.

::: callout-tip
### The real challenge in psychological science

- We are working together through four classes in **weeks 16-19**.
- These classes are designed to enable you to revise and put into practice some of the key ideas and skills you have been developing in the first year research methods modules *PSYC121, PSYC123 and PSYC124*.
- We have been doing this in the context of a live research project with potential real world impacts: the **Clearly Understood** project.
:::

In the week 19 class, we will face *the real challenge in psychological science*, in the context of the theory, methods, and data of the **Clearly Understood** project. This challenge results from the facts that:

1.  People vary
2.  Effects vary
3.  Results vary

- **Because** the effects of interventions, treatments or experimental manipulations can be expected to vary, 
- We **must** build the ability to investigate **how** those effects vary.

We will locate this ability in the capacity to work with **interaction effects** using the tools you have already been learning to apply within the linear model approach to analysis.

::: callout-tip
Our aim in these classes is to contribute new findings from the data we collect together.

- **That time is now.**

We will be working now with a large dataset that incorporates the responses from *PSYC122* students.
:::

### Our learning goals

In Week 19, we aim to extend your skills in *analyzing* and in *visualizing* psychological data.

We will use linear models to estimate the association between predictors and outcomes in order to answer our research questions.

::: callout-tip
**What is new** here is that we will examine how the effects that interest us can be different in different groups of people, in different contexts, or for different values of a third variable

- **If** Psychologists think that the shape or size of an effect is different in different groups of people, or in different contexts, or for different values of a third variable (a *moderator)*
- **Then** typically we are thinking about an **interaction** effect.
:::

When we work with linear models, as you have seen, we usually need to report:  

- information about the model we specify, identifying *all* predictors;
- our evaluation of whether the effects of one or more predictors are significant;
- model fit statistics (`F, R-squared`) as well as coefficient estimates;
- and descriptions of the impact of predictors.

Usually, in describing the impacts of predictors, we are required to communicate:

- the **direction** of the effect -- do values of the outcome variable *increase* or *decrease* given increasing values of the predictor?
- the **size** of the effect -- *how much* do values of the outcome variable *increase* or *decrease* given increasing values of the predictor?

::: callout-tip
In examining **interactions**, we now have to consider how we should analyse, visualize and talk about *how* effects may vary:

- Is the shape or size of an effect different in different groups of people, in different contexts, or for different values of a third variable (a *moderator*)?

We answer questions like these by building linear models that include **interaction effects**.
:::

## Lectures {#sec-wk19-lectures}

::: callout-tip
Before you go on to the activities in @sec-wk19-lab-activities, **watch** the lectures: `Week 19: Linear models – interactions and variation in effects`
:::

The lecture for this week is presented in four short parts.
You can view video recordings of the lectures using Panopto, by clicking on the video images shown following.

- Anybody who has the link should be able to view the video.

1. **Introduction** (9 minutes): Overview of the classes; introduction to key ideas -- the real challenge in psychological science (people vary, effects vary, results vary) -- variation, uncertainty, and working with the limits of samples.

```{=html}
<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=00bbc559-eb90-4766-a209-b40100c7ac33&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=false&interactivity=all" style="border: 1px solid #464646; position: absolute; top: 0; left: 0; width: 100%; height: 100%; box-sizing: border-box;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="PSYC122-lm-interactions-1-of-4"></iframe>
```

2. **The Clearly Understood project -- questions and analyses** (11 minutes): Critical thinking -- what effects can vary, how can effects vary, theoretical reasons motivating why effects might vary.

```{=html}
<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=2aabd1ed-c276-454b-bef8-b40100cc98c9&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=false&interactivity=all" style="border: 1px solid #464646; position: absolute; top: 0; left: 0; width: 100%; height: 100%; box-sizing: border-box;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="PSYC122-lm-interactions-2-of-4"></iframe>
```

3. **Coding models with interactions** (19 minutes): Revision on linear models as methods to predict outcomes; coding linear models with multiple predictors; models that do or do not include interactions; identifying the key information in the results from models including interactions; and making sense of what we see.

```{=html}
<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=4767b039-415a-4878-b048-b40100d74247&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=false&interactivity=all" style="border: 1px solid #464646; position: absolute; top: 0; left: 0; width: 100%; height: 100%; box-sizing: border-box;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="PSYC122-lm-interactions-3-of-4"></iframe>
```

4. **Interactions between numeric variables and factors** (13 minutes): Working with factors, the difference between numeric variables and factors, as predictors in linear models; the effects of factors in linear models; coding, plotting, interpreting and evaluating the effects of interactions between numeric variables and factors. Summary of key lessons.

```{=html}
<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=8531b025-7f1e-4ec3-9f67-b40100dd88fd&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=false&interactivity=all" style="border: 1px solid #464646; position: absolute; top: 0; left: 0; width: 100%; height: 100%; box-sizing: border-box;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="PSYC122-lm-interactions-4-of-4"></iframe>
```

::: callout-tip
The slides presented in the videos can be downloaded either as a web page or as a Word document.
:::

- [The slides](data/week19/122-linear-model-perspectives.html) exactly as presented (21 MB).
- [The slides](data/week19/122-linear-model-perspectives-printable-edit.docx) converted to a Word .docx (11 MB). 

You can download the web page `.html` file and click on it to open it in any browser (e.g., Chrome, Edge or Safari). The slide images are high quality and there are a lot of them, so the file is quite big; it may take a few seconds to download.

You can download the `.docx` file and click on it to open it as a Word document that you can then edit. 
Converting the slides to a `.docx` distorts some images but the benefit of the conversion is that it makes it easier for you to add your notes.

::: callout-tip
To work with our lecture recordings:

- Watch the video parts right through.
- Use the printable versions of the slides to make notes.
- Try out the coding exercises in the how-to guide and the activity tasks or questions (@sec-wk19-lab-activities) to learn how to construct visualizations and do analyses.
:::

### The lectures have four main areas of focus {#sec-wk19-lectures-focus}

**1. The real challenge in psychological science**

We look at a -- maybe *the* real challenge in psychological science -- as we attempt to build a scientific understanding of what people think and do, and why they do it. We connect (1.) the fact that psychological and social processes show variability to (2.) the uncertainties we face when we try to draw insights from limited sample evidence.

::: callout-tip
What makes psychological science challenging:

1. People vary
2. Effects
3. Results vary
:::

**2. Critical thinking**

We have been working on answers to research questions in the context of the **Clearly Understood** project on the comprehension of health information. 

- What person attributes predict success in understanding?
- Can people accurately evaluate whether they correctly understand written health information?

So far, we have been looking at the impact (on understanding) of variation between individuals in attributes like vocabulary knowledge.

Now, we need to consider the fact that the effect of one variable (a predictor like vocabulary knowledge) could itself vary.

::: callout-tip
We can develop our critical thinking skills by considering:

1. What effects can vary
2. How effects can vary

We discuss the kinds of theoretical assumptions that may lead Psychologists to hypothesize interaction effects.
:::

**3. Coding models with interactions**

We build on our understanding that linear models can be used to predict variation in psychological outcomes like observations of behaviours or scores on tests of person attributes. We build on our development of skills in coding linear models with multiple predictors to examine how we should code models incorporating interaction effects.

::: callout-tip
We build skills by looking at:

- The kind of coding we should use;
- The key information in results, where to find it, and how to interpret it;
- And the different ways in which we can present and interpret interactions.
:::

**4. Interactions between numeric variables and factors**

The effect on an outcome that is due to a numeric variable (like reading strategy) can differ depending on variation in another numeric variable (like vocabulary knowledge). But an effect can also differ depending on differences between groups or contexts. We often test or estimate the effects of differences between groups or contexts using categorical variables -- factors. This means we need to consolidate what we understand about working with factors in linear models. And then we need to build on this understanding by looking at interactions between numeric variables and factors.

::: callout-tip
We will learn about:

- How we code for the effects of factors in linear models;
- How we identify and interpret key information about factor effects;
- How we code models to estimate the effects of interactions between numeric variables and factors.
:::

We end the class by looking at the language and style of reporting results when we present our findings from linear model including an interaction effect.

## Reading: Links to other classes

We do not provide further reading for this class but you will find it helpful to revise some of the key ideas you have been learning about in PSYC122 and in other modules.

- The lectures in *PSYC122* on linear models. 
- The lectures in *PSYC123* on: the scientific method; reliability and validity; experimental design; and hypothesis testing.

## Pre-lab activities {#sec-wk19-prelab-activities}

### Pre-lab activity 1 {#sec-wk19-prelab-activities-1}

In weeks 16-19, we have been working together on a research project to investigate how people vary in their response to written health information or health advice.

Completing the project involves collecting responses from *PSYC122* students: **you**.

To enter your responses, we invite you to complete a short survey.

Complete the survey by clicking on the link [here](https://lancasteruni.eu.qualtrics.com/jfe/form/SV_8figdBSMYe4V2bI)

If you want more information about the project, we invite you to read the pre-registered research plan for the *PSYC122 health advice* research project.

[Read the project pre-registration](https://osf.io/p6fsc/)

## Lab activities {#sec-wk19-lab-activities}

### Introduction {#sec-wk19-lab-activities-1-introduction}

We will do our practical lab work in the context of the **Clearly Understood** project.

- Our focus is on what makes it easy or difficult for people to understand written health information.

::: callout-important
In Week 19, we aim to answer the research questions:

- What person attributes predict success in understanding health information?
- How can the effects of person attributes vary, in different groups of people, or for different values of a third variable (a *moderator*)?
:::

PSYC122 students contributed their responses to a survey we have been using to collect data to find answers to these questions.

In our practical work, we will be examining a dataset combining:

1. Data from two studies of the response of adults from a UK national sample to written health information

- `study-one-general-participants.csv`
- `study-two-general-participants.csv`

2. Data from PSYC122 students.

These data have been joined together to create a dataset of over 400 observations. This is because it often requires a lot of evidence to be able to draw precise or accurate conclusions about potential interaction effects.

### Get ready {#sec-wk19-prelab-activities-ready}

#### Download the data

Click on the link: [122_week19_for_students.zip](data/week19/122_week19_for_students.zip) to download the data files folder. Then upload the contents to the new folder you created in RStudio Server.

The downloadable .zip folder includes the data file:

- `all.data.csv`

and the R Markdown `.Rmd`:

- `PSYC122-w19-how-to.Rmd`

If you can't upload these files to the server -- this affects some students -- you can use some code to get R to do it for you: uncover the code box below to reveal the code to do this.

:::{.callout-tip collapse="true"}
## Code
- You can use the code below to directly download the file you need in this lab activity to the server.
- Remember that you can copy the code to your clipboard by clicking on the 'clipboard' in the top right corner.

Get the `all.data.csv` data:

```{r eval=FALSE}
download.file("https://github.com/lu-psy-r/statistics_for_psychologists/blob/main/PSYC122/data/week19/study-one-general-participants.csv?raw=true", destfile = "study-one-general-participants.csv")
```
::: 

If you need to get the `how-to` guide, you can do that by clicking on the link in the box below. 

- **Warning**: if you click on the link, the how-to guide might be shown to you (in some browsers) as a new webpage tab.

:::{.callout-tip collapse="true"}
## Link to the how-to guide
Get the `PSYC122-w19-how-to.Rmd` how-to guide

Click on the link: [PSYC122-w19-how-to.Rmd](data/week19/PSYC122-w19-how-to.Rmd) to download the `PSYC122-w19-how-to.Rmd` guide.
:::

#### Check: What is in the data file? {#sec-wk19-data-summary}

Each of the data files we will work with has a similar structure, as you can see in this extract.

```{r}
#| label: headcheck-wide
#| echo: false

target <- c('studyone.10', 'studytwo.1', 'R_8jSZdRgu4oq4QyD')

all.data %>%
  filter(participant_ID %in% target) %>%
  as.data.frame() %>%
  kable() %>%
  kable_styling()
```

You can use the *scroll bar* at the bottom of the data window to view different columns.

You can see the columns:

- `participant_ID` participant code;
- `mean.acc` average accuracy of response to questions testing understanding of health guidance (varies between 0-1);
- `mean.self` average self-rated accuracy of understanding of health guidance (varies between 1-9);
- `study` variable coding for what study the data were collected in;
- `AGE` age in years;
- `HLVA` health literacy test score (varies between 1-16);
- `SHIPLEY` vocabulary knowledge test score (varies between 0-40);
- `FACTOR3` reading strategy survey score (varies between 0-80);
- `GENDER` gender code (here, `Female, Male`);
- `EDUCATION` education level code (here, `Further, Higher`);
- `ETHNICITY` ethnicity (Office National Statistics categories) code (here, `Asian, Black, Mixed, White, Other`).

### Lab activity 1: Work with the `How-to` guide {#sec-wk19-lab-activities-1}

The `how-to` guide comprises an .Rmd file:

- `PSYC122-w19-how-to.Rmd`

It is full of advice and example code.

The code in the `how-to` guide was written to work with:

- `all.data.csv`

::: callout-tip
We show you how to do everything you need to do in the lab activity (@sec-wk19-lab-activities-2, shown following) in the `how-to` guide.

- Start by looking at the `how-to` guide to understand what steps you need to follow in the lab activity.
:::

We will take things *step-by-step*.

We split .Rmd scripts by steps, tasks and questions:

- different steps for different phases of the analysis workflow;
- different tasks for different things you need to do;
- different questions to examine different ideas or coding challenges.

::: callout-tip
- Make sure you start at the top of the `.Rmd` file and work in order through each task.
- Complete each task before you move on to the next task.
:::

In the activity @sec-wk19-lab-activities-2, we are going to work through a sequence of steps and tasks that mirrors the sequence you find in the `how-to` guide.

- There is a little bit of variation, comparing the later steps in the `how-to` guide and the steps in @sec-wk19-lab-activities-2, but that variation is designed to help you with your learning.

::: callout-tip
- Notice that we are gradually building up our skills: consolidating what we know; revising important learning; and extending ourselves to acquire new skills.
:::

**Step 1: Set-up**

1. Empty the R environment -- using `rm(list=ls())`
2. Load relevant libraries -- using `library()`

**Step 2: Load the data**

3. Read in the data file -- using `read_csv()`
4. Inspect the data -- using `summary()`

**Step 3: Work with different kinds of variables**

5. Visualize the distribution of a numeric variable -- using `geom_histogram()`
6. Identify the kind or class of a variable -- using functions like `is.factor()` or `is.numeric()` or `is.character()` -- the `is._()` family of functions
6. Change the kind or class of a variable -- using functions like `as.factor()` or `as.numeric()` or `as.character()` -- the `as._()` family of functions

**Step 4: Draw plots to examine associations between variables and to visualize interactions**

7. Draw scatterplots to compare the potential association between two variables -- using `geom_point()` 
7. Use `facet_wrap()` to examine or to show the association may vary for different values of a third variable (numeric or factor)
8. Use `cut_number()` to split a dataset, in rows (observations), by values of a numeric variable

**Step 5: Use linear models with multiple predictors, to estimate the effects of factors as well as numeric variables**

9. Fit a linear model including numeric variables as predictors
10. Fit a linear model including a numeric variable and a factor as predictors 

**Step 6: Use linear models with multiple predictors, including interaction effects**

11. Center numeric variables before using them as predictors
12. Specify models to include interaction effects: interactions between two numeric predictor variables -- using the `*` operator
13. Specify models to include interaction effects: interactions between a numeric predictor variable and a factor predictor variable -- using the `*` operator

::: callout-tip
If you are unsure about what you need to do, look at the advice in `PSYC122-w19-how-to.Rmd` on how to do the tasks, with examples on how to write the code.
:::

You will see that you can match a task in the activity @sec-wk19-lab-activities-2 to the same task in the `how-to` guide.

- The `how-to` shows you what function you need and how you should write the function code.

This process of adapting demonstration code is a process critical to data literacy and to effective problem solving in modern psychological science.

::: callout-warning
Don't forget: You will need to change the names of the variables to complete the tasks in @sec-wk19-lab-activities-2.
:::

### Lab activity 2 {#sec-wk19-lab-activities-2}

### OK: now let's do it!

In the following, we will guide you through the tasks and questions step by step.

::: callout-tip
1. We will not *at first* give you the answers to questions about the data or about the results of analyses.
2. An answers version of the workbook will be provided after the last lab session (check the answers then in @sec-wk19-lab-activities-answers) so that you can check whether your independent work has been correct.
:::

#### Questions {#sec-wk19-lab-activities-2-questions}

::: callout-warning
**Students have told us** that it would be helpful to your learning if we reduce the information in the hints we provide you. **We have done this** in Week 19.

The motivation for doing this is:

1. It will require you to do more *active* thinking to complete tasks or answer questions;
2. Thus, you can check to see how your learning is developing -- can you do the tasks, given what you know now?
3. Plus, psychological research shows that *active* thinking is better for understanding and for learning.

Where we do give you hints, we will sometimes replace the correct bit of code with a place-holder: `...`

- Your task will therefore be to replace the place-holder `...` with the correct bit of code or the correct data-set or variable name.
:::

#### Step 1: Set-up

To begin, we set up our environment in R.

##### Task 1 -- Run code to empty the R environment

:::{.callout-tip collapse="true"}
## Code
```{r}
#| eval: false
rm(list=ls())
```
:::

##### Task 2 -- Run code to load relevant libraries

Notice that in Week 19, we need to work with the libraries `ggeffects` and `tidyverse`.
Use the `library()` function to make these libraries available to you.

:::{.callout-tip collapse="true"}
## Code
```{r}
library("ggeffects")
library("tidyverse")
```
:::

#### Step 2: Load the data

##### Task 3 -- Read in the data file we will be using

The data files is called:

- `all-data.csv`

Use the `read_csv()` function to read the data files into R:

:::{.callout-tip collapse="true"}
## Code
```{r}
all.data <- read_csv("all.data.csv")  
```
:::

When you read the data files in, give the data object you create a meaningful name 
e.g. `all.data`.

##### Task 4 -- Inspect the data file

Use the `summary()` function to take a look at the dataset.

:::{.callout-tip collapse="true"}
## Code
```{r}
summary(all.data)
```
:::

#### Step 3: Work with different kinds of variables

It is an important skill in psychological data analysis to be able to: 

(1.) identify the different kinds of data variables we are working with;
(2.) adapt our methods depending on differences in kind (variable `Class` or type).

We learn about these skills, next, in exercises that combine some revision with
some new moves.

::: callout-tip
## Revise: consolidate what you know
:::

##### Task 5 -- Use summaries or plots to examine variables

##### Questions: Task 5

>**Q.1.** What is the median of the `FACTOR3` variable in the dataset?

<!-- - A.1. You can use the `summary()` function to calculate that the median `FACTOR3` score is 50.00 -->

>**Q.2.** Draw a histogram of the `FACTOR3` distribution.

<!-- - A.2. You can write the code using the same code structure you have been shown e.g. in `PSYC122-w19-how-to.Rmd` but changing data-set and variable names to get the right plots: -->

```{r}
ggplot(data = all.data, aes(x = FACTOR3)) +
  geom_histogram(binwidth = 1) +
  theme_bw() +
  labs(x = "Reading strategy (FACTOR3)", y = "frequency count")
```

>**Q.3.** Examine the reading strategy (`FACTOR3`) scores distribution: what does it tell you about the variation in the reading strategies of the participants in the sample?

<!-- - A.3. The histogram shows that most participants present scores varying somewhere between 40-60 but there is a "long tail" of smaller numbers of participants with low or very low scores on the reading strategy (`FACTOR3`) measure. -->

::: callout-tip
## Extend: make some new moves
:::

We can tell from the summary and the distribution that the `FACTOR3` variable consists of numbers. 

Another way to identify the kind or class of a variable is by using a check question:

```{r}
is.factor(all.data$FACTOR3)
is.numeric(all.data$FACTOR3)
```

The `is._()` family of functions act like identity checks: what is this variable? 
These functions work like questions with TRUE or FALSE answers.

- Q. Is this variable a factor? A. `[TRUE]` or `[FALSE]` 
- Q. Is this variable numeric? A. `[TRUE]` or `[FALSE]

##### Task 6 -- Identify (and change) what kind of variable -- numeric or factor -- a variable is

> Hint: Task 6 -- The `is._()` functions are available to ask what kind a variable is
> Hint: Task 6 -- The `as._()` functions can be used to change variable types

>**Q.4.** What kind of variable is `GENDER`?

<!-- - A.4. Use the same functions but change the variable name: -->

<!-- ```{r} -->
<!-- is.factor(all.data$GENDER) -->
<!-- is.numeric(all.data$GENDER) -->
<!-- ``` -->

>**Q.5.** What do the `is._()` functions tell us?

<!-- - A.5. The function calls tell us that `GENDER` is not a factor and is not numeric. -->

>**Q.6.** If you look at the `summary()` results, what does it tell you the `Class` of `GENDER` is?

<!-- - A.6. The summary shows that `GENDER` is identified as `Class: character`.  -->

<!-- You can check this by doing the `is._()` test again. -->

<!-- ```{r} -->
<!-- is.character(all.data$GENDER) -->
<!-- ``` -->

>**Q.7.** How do you change the class or type of the variable `GENDER`? Change the variable's type (`Class`) into a factor.

<!-- - A.7. The `as._()` functions can be used to change variable type, as you have done -->
<!-- previously (e.g., in week 16). -->

<!-- ```{r} -->
<!-- all.data$GENDER <- as.factor(all.data$GENDER) -->
<!-- ``` -->

>**Q.8.** What does a `summary()` tell us about the self-reported `GENDER` of participants in the `all.data` sample?

<!-- ```{r} -->
<!-- summary(all.data) -->
<!-- ```  -->

<!-- - A.8. The summary will show how many different participants report themselves -->
<!-- as belonging to one of the different `GENDER` self-identification groups in the corresponding survey question. -->

<!-- You should see counts of the numbers of people in different `GENDER` level  -->
<!-- groups in the total survey sample: -->

<!-- GENDER -->
<!-- Female:323 -->
<!-- Male  :155 -->

<!-- We can see that 323 participants identified themselves as `Female` in the survey. -->

<!-- We can see that 155 participants identified themselves as `Male` in the survey. -->

##### Gender identity representation in survey samples

We note, here, that in addition to the numbers shown in the summary, six participants in the survey data entered the `Non-Binary` identification response option, and two participants entered the `Prefer not to say` option, in response to the gender identification question in the survey.

In the present exercises, we are looking at the way in which the effect on outcomes of variables -- like variation in health literacy (`HLVA`) or reading strategy (`FACTOR3`) score -- vary for different values of a categorical variable (like `GENDER`). These kinds of interactions become more complicated to analyse where there are more than two categories or levels in a factor, especially when the numbers of participants in one or more categories or levels is relatively small. 

We think it is important to ensure representation of the responses and experiences of the full range of diversity among the populations we study. In future work, we hope to have collected sufficient numbers of observations from groups of participants like those entering responses like `Non-Binary` or `Prefer not to say` in order to be able to incorporate all possible between-group comparisons in analyses.

For now, we enter this explanation as a promise to to examine more representative samples in future work. Here is a link, for those who are interested, to reports of analyses of data from a survey conducted by a leading research organization (Pew Research Center). Follow the link to "Methodology" for information on how this organization exemplify good practice in obtaining representative samples.

<https://www.pewresearch.org/social-trends/2025/05/29/the-experiences-of-lgbtq-americans-today/>

#### Step 4: Now draw plots to examine associations between variables and to visualize interactions

Here, we are going to build on previous work to use plots to:

(1.) Examine the associations between pairs of variables, an outcome and a predictor
(2.) Examine how the association between two variables can be different for
different levels of a third variable

Scenario (2.) comes up if we are thinking about or working with *interaction* effects.

::: callout-tip
## Consolidation: practice to strengthen skills
:::

##### Task 7 -- Create a scatterplot to examine the association between two numeric variables

> Hint: Task 7 -- We are working with `geom_point()` and you need x and y aesthetic mappings.

Here, we can look at the potential association between variation in reading strategy (using the `FACTOR3` measure) and accuracy of understanding of health information (`mean.acc`).

Code and run a chunk of code to make the plot.

:::{.callout-tip collapse="true"}
## Code
```{r}
ggplot(data = all.data, aes(x = FACTOR3, y = mean.acc)) +
  geom_point() +
  theme_bw() +
  labs(y = "Accuracy of understanding (mean.acc)", 
       x = "Reading strategy (FACTOR3)") +
  xlim(0, 65) + ylim(0, 1)
```

This plot shows:

- the possible association between x-axis variable `FACTOR3` and y-axis 
variable `mean.acc`.

The plot code moves through the following steps:  

1. `ggplot(...)` make a plot;
2. `ggplot(data = all.data, ...)` with the `all.data` dataset;
3. `ggplot(...aes(x = FACTOR3, y = mean.acc))` using two aesthetic mappings

- `x = FACTOR3` map `FACTOR3` values to x-axis (horizontal, left to right) 
positions;
- `y = mean.acc` map `mean.acc` values to y-axis (vertical, bottom to top) 
positions;

4. `geom_point()` show the mappings as points.
5. `theme_bw()` changes the theme to a white background.
6. `labs(y = "Accuracy of understanding (mean.acc)", x = "Reading strategy (FACTOR3)")` changes axis labels.
7. `xlim(0, 65) + ylim(0, 1)` changes axis limits.
:::

>**Q.9.** What do you notice about the distribution of `mean.acc` scores at
increasing values of reading strategy (`FACTOR3`) scores?

<!-- - A.9. The scatterplot shows that, for increasing reading strategy (`FACTOR3`) scores, there is an associated rise in accuracy of understanding (`mean.acc`) scores. -->

<!-- - There is extra credit for noticing that variation in outcome and predictor variable values is not tightly coupled: it is possible to have low reading strategy (`FACTOR3`) scores and high accuracy of understanding (`mean.acc`). -->

::: callout-tip
## Introduce: make some new moves
:::

In the next sequence of exercises, we are going to take the `all.data` dataset and split it into parts (sub-sets) in different ways.

- We split the dataset vertically, into different sets of rows or observations.

##### Why are we learning how to do this?
  
The capacity to do this kind of split-analyse operation is *extremely* useful.

We are going to do this so that we can examine how the effects of interest to us
(here, associations) can vary between different groups of people.

Remember, observations about different people are on different rows in `all.data`.

- We are going to focus on an association between two column variables (`mean.acc`, `FACTOR3`).
- We are going to identify different groups (sub-sets) of observations using values on a third variable.

##### Task 8 -- Create a scatterplot to examine the association between two numeric variables, showing how the association is different for different values of a third variable

> Hint: Task 8 -- We are working with `geom_point()` again.

> Hint: Task 8 -- We can show how the association between two variables (`mean.acc`, `FACTOR3`) is different for different values of a third, categorical (factor), variable (`GENDER`) by splitting the plot so that we see the association for different levels of `GENDER`.

Here, we are splitting the dataset into:

- observation data from people with just `GENDER` level `Female`
- observation data from people with just `GENDER` level `Male`

and we are then producing a different scatterplot for each sub-set of data produced by the split.

Run the following chunk of code to make the plot.

:::{.callout-tip collapse="true"}
## Code
```{r}
ggplot(data = all.data, aes(x = FACTOR3, y = mean.acc)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  theme_bw() +
  labs(y = "Accuracy of understanding (mean.acc)", 
       x = "Reading strategy (FACTOR3)") +
  xlim(0, 65) + ylim(0, 1) +
  facet_wrap(~ GENDER)
```

This plot shows:

- the possible association between x-axis variable `FACTOR3` and y-axis 
variable `mean.acc`
- for different groups of people -- people with self-reported `GENDER` responses of `Female` or `Male`.

The plot code moves through steps we have seen before. What is new here is this bit:

`facet_wrap(~ GENDER)`

The function `facet_wrap()` splits the dataset into sub-sets (different parts): different sets of rows, where different sets are defined according to different values of the variable identified inside the brackets `(~ GENDER)`.

Here, we are asking for different scatterplots for the dataset split by whether
participants are coded as `Female` or `Male` for this survey.

You can see a general guide to the function here:

<https://ggplot2.tidyverse.org/reference/facet_wrap.html>
:::

Now use the plots to answer the following question.

>**Q.10.** What do you notice about the distribution of `mean.acc` scores at
increasing values of reading strategy (`FACTOR3`) scores, for different sub-sets of the data: for people with `Female` or `Male` `GENDER` coding?

<!-- - A.10. You can see that the association between `mean.acc` and `FACTOR3` scores appears to be *mostly* similar for different `GENDER` codes. -->

<!-- - The distribution of scores is different: we have more `Female` participants; their scores vary more widely. -->
<!-- - But the trend is similar overall: the trend line representing the association between `mean.acc` and `FACTOR3` looks similar such that for both sub-groups in the sample, increasing `FACTOR3` scores are associated with higher `mean.acc` scores. -->

> Hint: Task 8 -- We can show how the association between two variables (`mean.acc`, `FACTOR3`) is different for different values of a third numeric variable (`HLVA`) by splitting the dataset

Here, we are first creating a new variable to code for (distinguish between) different sub-sets of observations (different rows in the dataset) based on `HLVA` scores, and then second drawing different plots based on different subsets.

We do this in two steps, as follows.

First, run a chunk of code to divide (`cut`) the dataset into parts.

:::{.callout-tip collapse="true"}
```{r}
all.data$HLVA_splits <- cut_number(all.data$HLVA, 3)
```

The code works bit-by-bit like this:

1. `all.data$HLVA_splits <-` create a new variable, `HLVA_splits` and add it to the dataset `all.data` given the work done by the bit of code on the right of the arrow `<-`.

On the right of the arrow `<-` ...

2. `cut_number(all.data$HLVA, 3)` uses the `cut_number(...)` to divide the dataset observations into three sets based on the values in the `all.data$HLVA` variable named inside the brackets.

The function works by ordering all the observations (the rows) in the dataset according to the values in the named variable `all.data$HLVA`, then splitting the dataset (here, into three sub-sets) based on values in that named variable.
:::

Given this code, we are going to get three sub-sets of the data, observations corresponding to (1.) low (2.) mid or (3.) high `HLVA` values.

- We split the data into data about people with:

(1.) low `HLVA` values, scores between 2-8 on the `HLVA` test;
(2.) mid `HLVA` values, scores between 8-10 on the `HLVA` test;
(3.) high `HLVA` values, scores between 10-14 on the `HLVA` test.

You can see a guide to the key function here:

<https://ggplot2.tidyverse.org/reference/cut_interval.html>

Note that *where* we split the data (what ranges of scores we use) will depend on the sample.

Second, we draw a plot to examine if the association between two variables (`mean.acc`, `HLVA`) is different for different values of a third variable (`SHIPLEY`) by splitting the data

Run the following chunk of code to make the plot.

:::{.callout-tip collapse="true"}
## Code
```{r}
ggplot(data = all.data, aes(x = FACTOR3, y = mean.acc)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_wrap(~ HLVA_splits, nrow = 1, labeller = label_both) +
  labs(y = "Accuracy of understanding (mean.acc)", 
       x = "Reading strategy (FACTOR3)") +
  theme_bw() 
```

The plot code moves through the scatterplot production steps we have seen before.  

The bit that is new is here:

`facet_wrap(~ HLVA_splits, nrow = 1, labeller = label_both)` which is included to use the data split, constructed earlier.

This addition allows us to show:

- the nature of the association between two variables (`mean.acc`, `FACTOR3`) 
- for different values of the third variable (`HLVA`)
- by splitting the data into three sub-sets according to `HLVA` score (2-8 low, 8-10 mid, 10-14 high).
:::

Now use the plots to answer the following question.

>**Q.11.** What do you notice about the distribution of `mean.acc` scores at
increasing values of reading strategy (`FACTOR3`) scores, for different sub-sets of the data: for (1.) low (2.) mid (3.) high `HLVA` values?

<!-- - A.11. You can see that the association between `mean.acc` and `FACTOR3` scores appears to vary (the slope differs) given different values of the third variable (`HLVA`). -->

>**Q.12.** What do you think the values in the grey labels at the top of the facets (the plot panels) tell us?

<!-- - A.12. The values tell us what values (e.g., `2,8`) are used to identify the range of `HLVA` scores that have been used to define the sub-set of observations for which the plot below shows the association between `mean.acc` and `FACTOR3` scores. -->

#### Step 5: Use linear models with multiple predictors, to estimate the effects of factors as well as numeric variables

::: callout-tip
## Consolidation: practice to strengthen skills
:::

As we saw in weeks 17 and 18, we can use linear models to predict outcome variables. 

- Linear models can include numeric variables (like `FACTOR3` or `HLVA`) as predictors. 
- Linear models can also include categorical variables or factors (like `GENDER` or `EDUCATION`) as predictors.

##### Task 9 -- First, fit a linear model including just numeric variables as predictors

> Hint: Task 9 -- We use the `lm()` function to fit a model with `mean.acc` as the outcome and `FACTOR3` and `HLVA` as predictors.

Note that in this model:

- `FACTOR3` is a measure of reading strategy -- numeric scores correspond to a test of the relative effectiveness of the strategies participants use to read and understand written information;
- `HLVA` is a measure of health literacy -- numeric scores correspond to a test of knowledge of health-related words.

:::{.callout-tip collapse="true"}
## Code
```{r}
model <- lm(mean.acc ~ FACTOR3 + HLVA, 
            data = all.data)

summary(model)
```
:::

If you look at the model summary you can answer the following questions.

>**Q.13.** What is the estimate for the coefficient of the effect of the predictor, `FACTOR3` (reading strategy)?

<!-- - A.13. 0.0023489 -->

>**Q.14.** Is the effect significant?

<!-- - A.14. It is significant. Because `Pr(>|t|)` equals `0.003` for the `Estimate` of the effect of `FACTOR3` (reading strategy): this means that  p < .05, and so the effect is significant. -->

>**Q.15.** What are the values for t and p for the significance test for the 
estimated coefficient of the effect of `HLVA` (health literacy)?

<!-- - A.15. t = 8.861, p = <2e-16 -->

:::{.callout-tip collapse="true"}
## Do you know how to intepret numbers like `<2e-16`?

If you are unsure you can find out about scientific notation here:

<https://www.calculatorsoup.com/calculators/math/scientific-notation-converter.php>
:::

>**Q.16.** What do you conclude are the effects of the `FACTOR3` (reading strategy) and `HLVA` (health literacy) variables, as predictors of outcome `mean.acc` (accuracy of understanding of health information)?

<!-- - A.16. The model slope estimates suggests that both predictors are significant, and that as both `FACTOR3` (reading strategy) and `HLVA` (health literacy) increase, so `mean.acc` scores are predicted to increase also. -->

##### Task 10 -- Second, fit a linear model including a numeric variable and a factor as predictors

> Hint: Task 10 -- We use the `lm()` function to fit a model with `mean.acc` as the outcome and `FACTOR3` and `GENDER` as predictors.

Note that in this model:

- `FACTOR3` is a measure of reading strategy -- numeric scores correspond to a test of strategy awareness
- `GENDER` is a self-report measure of gender identity

:::{.callout-tip collapse="true"}
```{r}
model <- lm(mean.acc ~ FACTOR3 + GENDER, 
            data = all.data)

summary(model)
```
:::

If you look at the model summary you can answer the following questions.

>**Q.17.** What is the estimate for the coefficient of the effect of the predictor, `FACTOR3` (strategy)?

<!-- - A.17. 0.0040658 -->

>**Q.18.** Is the effect significant?

<!-- - A.18. It is significant. Because `Pr(>|t|)` equals `1.1e-06` for the `Estimate` of the effect of `FACTOR3` (strategy): this means that  p < .05, and so the effect is significant. -->

>**Q.19.** What are the values for t and p for the significance test for the 
estimated coefficient of the effect of `GENDER` (level, `Male` compared to `Female`)?

<!-- - A.19. t = -0.901, p = 0.368  -->

>**Q.20.** Is the effect of `GENDER` significant?

<!-- - A.20. It is not significant. Because `Pr(>|t|)` equals `0.635` for the `Estimate` of the effect of `GENDER` this means that  p > .05, and so the effect is not significant.  -->

> Hint: Task 10 -- How do we know how to interpret estimates for the effects of factors? We discussed how R deals with factors in week 18, and discuss factors, and the effects of factors, in the week 19 lecture.

>**Q.21.** How should we interpret the coefficient estimate for the effect of `GENDER`?

<!-- - A.21. Though the effect is not significant, we can say that, on average, outcome `mean.acc` is -0.0108184 (slightly) lower for `Male` compared to `Female` coded people. -->

>**Q.22.** What do you conclude are the effects of the `FACTOR3` (strategy) and `GENDER` (`Female` compared to `Male`) variables, as predictors of outcome `mean.acc` (accuracy of understanding of health information)?

<!-- - A.22. The model slope estimates suggest that gender does not have a significant effect but that as `FACTOR3` (strategy) scores increase, so `mean.acc` scores are predicted to increase.  -->

#### Step 6: Use linear models with multiple predictors, including interaction effects

::: callout-tip
## Introduce: make some new moves
:::

As we saw through Task 8, it is possible that:

- the association between two variables (`mean.acc`, `FACTOR3`) can be different for different values of a third categorical (factor) variable (`GENDER`);
- the association between two variables (`mean.acc`, `FACTOR3`) can be different for different values of a third variable (`HLVA`)

While we can examine these possibilities using visualizations, we often need to conduct statistical analyses to test *interaction effects*, or the ways in which the association between two variables can be different for different levels of a third variable.

We can extend linear models to include terms that allow us to test or to estimate *interaction effects*.

##### Task 11 -- Center numeric variables before using them as predictors

> Hint: Task 11

When we include numeric variables (e.g., `FACTOR3, HLVA, SHIPLEY`) in linear models, it can cause problems of interpretation or of estimation if we include the variables as they first come to us in datasets (as *raw* variables).

- It often helps our analyses to *center* variables on their means.
- Centering a variable means calculating its mean, and then subtracting that mean from every value in the variable column.

For example, to center values of the `HLVA` variable, we create a new column `data$HLVA_centered` by first calculating the mean of the `data$HLVA` variable, and then subtracting that mean from every row value in `data$HLVA`.

:::{.callout-tip collapse="true"}
## Code
```{r}
all.data$HLVA_centered <- all.data$HLVA - mean(all.data$HLVA, na.rm = TRUE)
```
:::

>**Q.23.** Can you check what is going on when we center variables?

- Q.23. Hint: You can see what is happening when you run this code by first calculating the mean for `HLVA`

:::{.callout-tip collapse="true"}
## Code
```{r}
mean(all.data$HLVA, na.rm = TRUE)
```
:::

then looking at column values in the new `HLVA_centered` column, compared to the original `HLVA` column:

:::{.callout-tip collapse="true"}
## Code
```{r}
all.data %>%
  select(HLVA_centered, HLVA) %>%
  print(n = 20)
```
:::

What is the difference between the `HLVA` and `HLVA_centered` columns?

<!-- - A.23. You can see that every value in the `HLVA_centered` column equals the value, on the same row, in the `HLVA` column, minus the mean for `HLVA`. -->

>**Q.24.** Can you center the variable `FACTOR3`?

<!-- - A.24. You can use the same code to center `FACTOR3`: -->

<!-- ```{r} -->
<!-- all.data$FACTOR3_centered <- all.data$FACTOR3 - mean(all.data$FACTOR3, na.rm = TRUE) -->
<!-- ``` -->

>**Q.25.** Can you center the variable `SHIPLEY`?

<!-- - A.25. You can use the same code to center `SHIPLEY`: -->

<!-- ```{r} -->
<!-- all.data$SHIPLEY_centered <- all.data$SHIPLEY - mean(all.data$SHIPLEY, na.rm = TRUE) -->
<!-- ``` -->

##### Why are we learning how to do this?

It is common in data analysis to want to transform variables before using them in analyses.

- Here, we are transforming predictor variables by centering them on their mean values.

The practical reason for doing this transformation now is because if we use uncentered (raw) numeric variables as predictors in linear models that involve interactions the model can find it difficult to distinguish between the effects of those predictors and the effect of their interaction.

##### Task 12 -- Specify models to include interaction effects: interactions between two numeric predictor variables

> Hint: Task 12

We use `lm()` to specify models, as we have been doing. 

- What changes is that we now use the `*` operator in our model code to specify interaction effects.

For example, let's suppose that we want to examine the possibility that the effect of reading strategy (`FACTOR3`) on accuracy understanding (`mean.acc`) is different for different values of health literacy (`HLVA`).

- Here, we are considering the possibility that the association between reading strategy (`FACTOR3`) and accuracy of understanding (`mean.acc`) *varies*.
- We are considering the possibility that the the association between reading strategy (`FACTOR3`) and accuracy of understanding (`mean.acc`) is different for different groups of people.
- We are considering the possibility that the effect of reading strategy (`FACTOR3`) on accuracy of understanding (`mean.acc`) is different for people who differ, also, in health literacy (`HLVA`).

In examining the possibility that the effect of reading strategy (`FACTOR3`) on accuracy of understanding (`mean.acc`) is different for different values of health literacy (`HLVA`), we are examining a possible *interaction* between the effects of reading strategy (`FACTOR3`) and health literacy (`HLVA`).

::: callout-tip
## Let's go through this step-by-step
:::

First, consider: if we ignored the possibility of an interaction, we would fit a model with just the predictors: reading strategy (`FACTOR3`) and health literacy (`HLVA`).

We can fit the same model that we fit before (for Task 9).

- We vary the code a little, by using the centered variables we have just created:

:::{.callout-tip collapse="true"}
## Code
```{r}
model <- lm(mean.acc ~ FACTOR3_centered + HLVA_centered, 
            data = all.data)

summary(model)
```
:::

>**Q.26.** Can you compare the summary of results for this model with the model, using the same (but not centered) predictors, that you fitted for Task 9 -- compare the estimates, what do you see?

<!-- - A.26. If we compare the results of the models we can see that the estimates, t-test values and probability values are the same for `FACTOR3` compared to `FACTOR3_centered` or for `HLVA` compared to `HLVA_centered`.  -->

<!-- - This shows that centering predictors only changes the scaling of the predictor variables so that their centers are 0. -->

Second, now consider: we aim to test or estimate the interactions between the effects of reading strategy (`FACTOR3_centered`) and health literacy (`HLVA_centered`).

- We vary the code a little, again, this time by separating the predictors by `*` not by `+`:

:::{.callout-tip collapse="true"}
## Code
```{r}
model <- lm(mean.acc ~ FACTOR3_centered*HLVA_centered, 
            data = all.data)

summary(model)
```
:::

>**Q.27.** Can you compare the summary of results for this model with the results for the previous model (which uses the same predictors but without an interaction)?

- What do you see?

<!-- - A.27. If we compare the results of the models we can see that now the summary gives us estimates for the effects of: -->

<!-- - `FACTOR3_centered` -- this is the effect of `SHIPLEY_centered` when `HLVA_centered` is 0 -->
<!-- - `HLVA_centered` -- this is the effect of `HLVA_centered` when `FACTOR3_centered` is 0 -->
<!-- - `FACTOR3_centered*HLVA_centered` -- this is the interaction, and reflects the change in the effect of `FACTOR3_centered` on outcome `mean.acc`, given different values of `HLVA_centered`. -->

<!-- - Notice that the effect of the interaction is not significant but that the results suggest a very slight tendencey for the effect of reading strategy to be different (smaller) for increasing values of `HLVA_centered`. -->

##### What are we learning here?

The `*` symbol (the `*` operator) is used here to code for a model including the effects of the two variables separated by the symbol, here, as well as the effect of the interaction between these two variables.

### Task 13 -- Plot model predictions to help with interpretation of interaction effects

> Hint: Task 13 

How do we interpret interactions?

- It helps to plot model predictions, given the effects estimates that we derive from our sample using the linear model.

We can do this in a two-step sequence, similar to the sequence we trialled through weeks 17 and 18:

1. First fit the model

:::{.callout-tip collapse="true"}
## Code
```{r}
# -- first fit the model
model <- lm(mean.acc ~ FACTOR3_centered*HLVA_centered, 
            data = all.data)
```
:::

2. Second, use the model information to plot the predictions, given the effects

:::{.callout-tip collapse="true"}
## Code
```{r}
# -- second use model information to make a plot to show model predictions
plot_model(model, type = "pred", 
           terms = c("FACTOR3_centered", "HLVA_centered")) +
  theme_bw() +
  ylim(0, 1)
```
:::

>**Q.28.** Can you compare the summary of results for the model with the plot to develop an interpretation of the `FACTOR3_centered*HLVA_centered` interaction?

- What do you see?

<!-- - A.28. We can build an interpretation in steps: -->

<!-- - If we look at the coefficient estimate for the `FACTOR3_centered*HLVA_centered` interaction, we can see that it is negative (`-0.0005713`).  -->
<!-- - If we look at the plot of the model predictions, we can see that there are three slopes for the `FACTOR3_centered` effect, one each for different levels of `HLVA_centered`: for `HLVA_centered` = -2.19; `HLVA_centered` = 0; and `HLVA_centered` = 2.19. -->

<!-- - Notice that the steepness of the slope of the `FACTOR3_centered` effect decreases (the slope gets less steep) from `HLVA_centered` = -2.19 to `HLVA_centered` = 2.19. -->
<!-- - This tells us that the interaction (though not significant) operates so that the effect of reading strategy tends to be smaller for people with higher levels of health literacy. -->

<!-- - Returning to the fact that the coefficient estimate for the `FACTOR3_centered*HLVA_centered` is negative, we can now understand that it is negative because the interaction operates so that the slope of the `FACTOR3_centered` gets smaller as scores on `HLVA_centered` increase. -->

### Task 14 -- Specify models to include interaction effects: interactions between a numeric predictor variable and a factor predictor variable

> Hint: Task 14 -- We follow the same approach that we followed through Task 12.

Now consider: we aim to test or estimate the interactions between the effects of reading strategy (`FACTOR3`) and gender (`GENDER`).

- We can vary the code to change the predictor variables but the structure otherwise stays the same: how should we do this?

:::{.callout-tip collapse="true"}
## Code
```{r}
model <- lm(mean.acc ~ FACTOR3_centered*GENDER, 
            data = all.data)

summary(model)
```
:::

>**Q.29.** Can you identify what effects are shown in the model summary?

<!-- - A.29. We can see that now the summary gives us estimates for the effects of: -->

<!-- - `FACTOR3_centered` -- this is the effect of reading strategy (`FACTOR3`) when `GENDER` level is `Female` -->
<!-- - `GENDERMale` -- this is the effect of `GENDER` when `FACTOR3_centered` is 0.  -->

<!-- Note that the effect of `GENDER` is the estimated difference in outcome when `GENDER` level is `Female` (not shown, treated as the baseline) compared to when `GENDER` level is `Male`. -->

<!-- - `FACTOR3_centered*GENDER` -- this is the interaction, and reflects the change in the effect of `FACTOR3_centered` on outcome `mean.acc`, given different values of `GENDER`, here, when `GENDER` is `Male` instead of `Female`. -->

##### What are we learning here?

Remember from week 18:

Categorical variables or **factors** and reference levels.

- If you have a categorical variable like `GENDER` then when you use it in
an analysis, R will look at the different categories (called `levels) e.g., here,
`Female, Male` and it will pick one level to be the *reference* or baseline level. 
- The *reference* is the the level against which other levels are compared. 
- Here, the reference level is `Female` simply because, unless
you tell R otherwise, it picks the level with a category name that begins
earlier in the alphabet as the reference level.

##### How do we interpret interactions?

- It helps to plot the effects estimates, given the model.

We can do this in a two-step sequence, similar to the sequence we trialled through weeks 17 and 18:

1. First fit the model
2. Second, use the model information to plot the predictions, given the effects

:::{.callout-tip collapse="true"}
## Code
```{r}
# -- first fit the model
model <- lm(mean.acc ~ FACTOR3_centered*GENDER, 
            data = all.data)

# -- make plot
plot_model(model, type = "pred", 
           terms = c("FACTOR3_centered", "GENDER")) +
  theme_bw() +
  ylim(0, 1)
```
:::

>**Q.30.** Can you identify how the effects of reading strategy (`FACTOR3`) and gender (`GENDER`) interact, given the model results summary and this plot?

<!-- - A.30. The plot, and the estimate for the interaction `FACTOR3_centered:GENDERMale`, together show that the coefficient for the effect of reading strategy (`FACTOR3_centered`) is slightly larger (the slope is slightly steeper) when `GENDER` is `Male` instead of `Female. -->

<!-- - Notice that, for this sample, the interaction is not significant, suggesting that there is no difference between genders in the relative effectiveness of reading strategy. -->

#### You have now completed the Week 19 questions.

You have now extended your capacity to *think critically* about our data and our capacity to predict people and their behaviour.

::: callout-tip
- We have used PSYC122 students' responses to examine the robustness of evidence for potential answers to our research questions.
- Examining the ways that effects can vary between groups, contexts or the values of a moderator is a key element in the process of *building* insights in psychological science.
:::

## Answers {#sec-wk19-lab-activities-answers}

When you have completed all of the lab content, you may want to check your answers with our completed version of the script for this week.

::: callout-tip
The `.Rmd` script containing all code and all answers for each task and each question will be made available after the final lab session has taken place.

<!-- - You can download the script by clicking on the link:  [PSYC122-w19-workbook-answers.Rmd](data/week19/PSYC122-w19-workbook-answers.Rmd). -->
:::

We set out answers information on the Week 19 questions, below.

- We focus on the **Lab activity 2** questions where we ask you to interpret something or say something.
- We do not show questions where we have given example or target code in the foregoing lab activity @sec-wk19-lab-activities-2.

You can see all the code and all the answers in `PSYC122-w19-workbook-answers.Rmd`.

### Answers {#sec-wk19-lab-activities-answers-information}

::: callout-tip
Click on a box to reveal the answer.
:::

#### Questions {#sec-wk19-lab-activities-2-questions-answers}

>**Q.1.** What is the median of the `FACTOR3` variable in the dataset?

::: {.callout-note icon=false collapse="true"}
- A.1. You can use the `summary()` function to calculate that the median `FACTOR3` score is 50.00
:::

>**Q.2.** Draw a histogram of the `FACTOR3` distribution.

::: {.callout-note icon=false collapse="true"}
- A.2. You can write the code using the same code structure you have been shown e.g. in `PSYC122-w19-how-to.Rmd` but changing data-set and variable names to get the right plots:

```{r}
ggplot(data = all.data, aes(x = FACTOR3)) +
  geom_histogram(binwidth = 1) +
  theme_bw() +
  labs(x = "Reading strategy (FACTOR3)", y = "frequency count")
```
:::

>**Q.3.** Examine the reading strategy (`FACTOR3`) scores distribution: what does it tell you about the variation in the reading strategies of the participants in the sample?

::: {.callout-note icon=false collapse="true"}
- A.3. The histogram shows that most participants present scores varying somewhere between 40-60 but there is a "long tail" of smaller numbers of participants with low or very low scores on the reading strategy (`FACTOR3`) measure.
:::

>**Q.4.** What kind of variable is `GENDER`?

::: {.callout-note icon=false collapse="true"}
- A.4. Use the same functions but change the variable name:

```{r}
is.factor(all.data$GENDER)
is.numeric(all.data$GENDER)
```
:::

>**Q.5.** What do the `is._()` functions tell us?

::: {.callout-note icon=false collapse="true"}
- A.5. The function calls tell us that `GENDER` is not a factor and is not numeric.
:::

>**Q.6.** If you look at the `summary()` results, what does it tell you the `Class` of `GENDER` is?

::: {.callout-note icon=false collapse="true"}
- A.6. The summary shows that `GENDER` is identified as `Class: character`.

You can check this by doing the `is._()` test again.

```{r}
is.character(all.data$GENDER)
```
:::

>**Q.7.** How do you change the class or type of the variable `GENDER`? Change the variable's type (`Class`) into a factor.

::: {.callout-note icon=false collapse="true"}
- A.7. The `as._()` functions can be used to change variable type, as you have done
previously (e.g., in week 16).

```{r}
all.data$GENDER <- as.factor(all.data$GENDER)
```
:::

>**Q.8.** What does a `summary()` tell us about the self-reported `GENDER` of participants in the `all.data` sample?

::: {.callout-note icon=false collapse="true"}
```{r}
summary(all.data)
```

- A.8. The summary will show how many different participants report themselves
as belonging to one of the different `GENDER` self-identification groups in the corresponding survey question.

You should see counts of the numbers of people in different `GENDER` level
groups in the total survey sample:

GENDER
Female:323
Male  :155

We can see that 323 participants identified themselves as `Female` in the survey.

We can see that 155 participants identified themselves as `Male` in the survey.
:::

Create a scatterplot to examine the association between two numeric variables.

Here, we can look at the potential association between variation in reading strategy (using the `FACTOR3` measure) and accuracy of understanding of health information (`mean.acc`).

>**Q.9.** What do you notice about the distribution of `mean.acc` scores at
increasing values of reading strategy (`FACTOR3`) scores?

::: {.callout-note icon=false collapse="true"}
- A.9. The scatterplot shows that, for increasing reading strategy (`FACTOR3`) scores, there is an associated rise in accuracy of understanding (`mean.acc`) scores.

- There is extra credit for noticing that variation in outcome and predictor variable values is not tightly coupled: it is possible to have low reading strategy (`FACTOR3`) scores and high accuracy of understanding (`mean.acc`).
:::

Create a scatterplot to examine the association between two numeric variables, showing how the association is different for different values of a third variable.

We can show how the association between two variables (`mean.acc`, `FACTOR3`) is different for different values of a third, categorical (factor), variable (`GENDER`) by splitting the plot so that we see the association for different levels of `GENDER`.

>**Q.10.** What do you notice about the distribution of `mean.acc` scores at
increasing values of reading strategy (`FACTOR3`) scores, for different sub-sets of the data: for people with `Female` or `Male` `GENDER` coding?

::: {.callout-note icon=false collapse="true"}
- A.10. You can see that the association between `mean.acc` and `FACTOR3` scores appears to be *mostly* similar for different `GENDER` codes.

- The distribution of scores is different: we have more `Female` participants; their scores vary more widely.
- But the trend is similar overall: the trend line representing the association between `mean.acc` and `FACTOR3` looks similar such that for both sub-groups in the sample, increasing `FACTOR3` scores are associated with higher `mean.acc` scores.
:::

We can show how the association between two variables (`mean.acc`, `FACTOR3`) is different for different values of a third numeric variable (`HLVA`) by splitting the dataset.

Here, we are first creating a new variable to code for (distinguish between) different sub-sets of observations (different rows in the dataset) based on `HLVA` scores, and then second drawing different plots based on different subsets.

>**Q.11.** What do you notice about the distribution of `mean.acc` scores at
increasing values of reading strategy (`FACTOR3`) scores, for different sub-sets of the data: for (1.) low (2.) mid (3.) high `HLVA` values?

::: {.callout-note icon=false collapse="true"}
- A.11. You can see that the association between `mean.acc` and `FACTOR3` scores appears to vary (the slope differs) given different values of the third variable (`HLVA`).
:::

>**Q.12.** What do you think the values in the grey labels at the top of the facets (the plot panels) tell us?

::: {.callout-note icon=false collapse="true"}
- A.12. The values tell us what values (e.g., `2,8`) are used to identify the range of `HLVA` scores that have been used to define the sub-set of observations for which the plot below shows the association between `mean.acc` and `FACTOR3` scores.
:::

We use the `lm()` function to fit a model with `mean.acc` as the outcome and `FACTOR3` and `HLVA` as predictors.

>**Q.13.** What is the estimate for the coefficient of the effect of the predictor, `FACTOR3` (reading strategy)?

::: {.callout-note icon=false collapse="true"}
- A.13. 0.0023489
:::

>**Q.14.** Is the effect significant?

::: {.callout-note icon=false collapse="true"}
- A.14. It is significant. Because `Pr(>|t|)` equals `0.003` for the `Estimate` of the effect of `FACTOR3` (reading strategy): this means that  p < .05, and so the effect is significant.
:::

>**Q.15.** What are the values for t and p for the significance test for the 
estimated coefficient of the effect of `HLVA` (health literacy)?

::: {.callout-note icon=false collapse="true"}
- A.15. t = 8.861, p = <2e-16
:::

>**Q.16.** What do you conclude are the effects of the `FACTOR3` (reading strategy) and `HLVA` (health literacy) variables, as predictors of outcome `mean.acc` (accuracy of understanding of health information)?

::: {.callout-note icon=false collapse="true"}
- A.16. The model slope estimates suggests that both predictors are significant, and that as both `FACTOR3` (reading strategy) and `HLVA` (health literacy) increase, so `mean.acc` scores are predicted to increase also.
:::

We use the `lm()` function to fit a model with `mean.acc` as the outcome and `FACTOR3` and `GENDER` as predictors.

>**Q.17.** What is the estimate for the coefficient of the effect of the predictor, `FACTOR3` (strategy)?

::: {.callout-note icon=false collapse="true"}
- A.17. 0.0040658
:::

>**Q.18.** Is the effect significant?

::: {.callout-note icon=false collapse="true"}
- A.18. It is significant. Because `Pr(>|t|)` equals `1.1e-06` for the `Estimate` of the effect of `FACTOR3` (strategy): this means that  p < .05, and so the effect is significant.
:::

>**Q.19.** What are the values for t and p for the significance test for the 
estimated coefficient of the effect of `GENDER` (level, `Male` compared to `Female`)?

::: {.callout-note icon=false collapse="true"}
- A.19. t = -0.901, p = 0.368
:::

>**Q.20.** Is the effect of `GENDER` significant?

::: {.callout-note icon=false collapse="true"}
- A.20. It is not significant. Because `Pr(>|t|)` equals `0.635` for the `Estimate` of the effect of `GENDER` this means that  p > .05, and so the effect is not significant.
:::

>**Q.21.** How should we interpret the coefficient estimate for the effect of `GENDER`?

::: {.callout-note icon=false collapse="true"}
- A.21. Though the effect is not significant, we can say that, on average, outcome `mean.acc` is -0.0108184 (slightly) lower for `Male` compared to `Female` coded people.
:::

>**Q.22.** What do you conclude are the effects of the `FACTOR3` (strategy) and `GENDER` (`Female` compared to `Male`) variables, as predictors of outcome `mean.acc` (accuracy of understanding of health information)?

::: {.callout-note icon=false collapse="true"}
- A.22. The model slope estimates suggest that gender does not have a significant effect but that as `FACTOR3` (strategy) scores increase, so `mean.acc` scores are predicted to increase.
:::

We can extend linear models to include terms that allow us to test or to estimate *interaction effects*.

When we include numeric variables (e.g., `FACTOR3, HLVA, SHIPLEY`) in linear models, it can cause problems of interpretation or of estimation if we include the variables as they first come to us in datasets (as *raw* variables).

- It often helps our analyses to *center* variables on their means.
- Centering a variable means calculating its mean, and then subtracting that mean from every value in the variable column.

>**Q.23.** Can you check what is going on when we center variables?

- Q.23. Hint: You can see what is happening when you run this code by first calculating the mean for `HLVA`

:::{.callout-tip collapse="true"}
## Code
```{r}
mean(all.data$HLVA, na.rm = TRUE)
```
:::

then looking at column values in the new `HLVA_centered` column, compared to the original `HLVA` column:

:::{.callout-tip collapse="true"}
## Code
```{r}
all.data %>%
  select(HLVA_centered, HLVA) %>%
  print(n = 20)
```
:::

What is the difference between the `HLVA` and `HLVA_centered` columns?

::: {.callout-note icon=false collapse="true"}
- A.23. You can see that every value in the `HLVA_centered` column equals the value, on the same row, in the `HLVA` column, minus the mean for `HLVA`.
:::

>**Q.24.** Can you center the variable `FACTOR3`?

::: {.callout-note icon=false collapse="true"}
- A.24. You can use the same code to center `FACTOR3`:

```{r}
all.data$FACTOR3_centered <- all.data$FACTOR3 - mean(all.data$FACTOR3, na.rm = TRUE)
```
:::

>**Q.25.** Can you center the variable `SHIPLEY`?

::: {.callout-note icon=false collapse="true"}
- A.25. You can use the same code to center `SHIPLEY`:

```{r}
all.data$SHIPLEY_centered <- all.data$SHIPLEY - mean(all.data$SHIPLEY, na.rm = TRUE)
```
:::

Specify models to include interaction effects: interactions between two numeric predictor variables.

For example, let's suppose that we want to examine the possibility that the effect of reading strategy (`FACTOR3`) on accuracy understanding (`mean.acc`) is different for different values of health literacy (`HLVA`).

First, consider: if we ignored the possibility of an interaction, we would fit a model with just the predictors: reading strategy (`FACTOR3`) and health literacy (`HLVA`).

We can fit the same model that we fit before (for Task 9).

- We vary the code a little, by using the centered variables we have just created.

>**Q.26.** Can you compare the summary of results for this model with the model, using the same (but not centered) predictors, that you fitted for Task 9 -- compare the estimates, what do you see?

::: {.callout-note icon=false collapse="true"}
- A.26. If we compare the results of the models we can see that the estimates, t-test values and probability values are the same for `FACTOR3` compared to `FACTOR3_centered` or for `HLVA` compared to `HLVA_centered`.

- This shows that centering predictors only changes the scaling of the predictor variables so that their centers are 0.
:::

Second, now consider: we aim to test or estimate the interactions between the effects of reading strategy (`FACTOR3_centered`) and health literacy (`HLVA_centered`).

- We vary the code a little, again, this time by separating the predictors by `*` not by `+`.

>**Q.27.** Can you compare the summary of results for this model with the results for the previous model (which uses the same predictors but without an interaction)?

- What do you see?

::: {.callout-note icon=false collapse="true"}
- A.27. If we compare the results of the models we can see that now the summary gives us estimates for the effects of:

- `FACTOR3_centered` -- this is the effect of `SHIPLEY_centered` when `HLVA_centered` is 0
- `HLVA_centered` -- this is the effect of `HLVA_centered` when `FACTOR3_centered` is 0
- `FACTOR3_centered*HLVA_centered` -- this is the interaction, and reflects the change in the effect of `FACTOR3_centered` on outcome `mean.acc`, given different values of `HLVA_centered`.

- Notice that the effect of the interaction is not significant but that the results suggest a very slight tendencey for the effect of reading strategy to be different (smaller) for increasing values of `HLVA_centered`.
:::

Plot model predictions to help with interpretation of interaction effects.

- It helps to plot model predictions, given the effects estimates that we derive from our sample using the linear model.

>**Q.28.** Can you compare the summary of results for the model with the plot to develop an interpretation of the `FACTOR3_centered*HLVA_centered` interaction?

- What do you see?

::: {.callout-note icon=false collapse="true"}
- A.28. We can build an interpretation in steps:

- If we look at the coefficient estimate for the `FACTOR3_centered*HLVA_centered` interaction, we can see that it is negative (`-0.0005713`).
- If we look at the plot of the model predictions, we can see that there are three slopes for the `FACTOR3_centered` effect, one each for different levels of `HLVA_centered`: for `HLVA_centered` = -2.19; `HLVA_centered` = 0; and `HLVA_centered` = 2.19.

- Notice that the steepness of the slope of the `FACTOR3_centered` effect decreases (the slope gets less steep) from `HLVA_centered` = -2.19 to `HLVA_centered` = 2.19.
- This tells us that the interaction (though not significant) operates so that the effect of reading strategy tends to be smaller for people with higher levels of health literacy.

- Returning to the fact that the coefficient estimate for the `FACTOR3_centered*HLVA_centered` is negative, we can now understand that it is negative because the interaction operates so that the slope of the `FACTOR3_centered` gets smaller as scores on `HLVA_centered` increase.
:::

Specify models to include interaction effects: interactions between a numeric predictor variable and a factor predictor variable

Now consider: we aim to test or estimate the interactions between the effects of reading strategy (`FACTOR3`) and gender (`GENDER`).

- We can vary the code to change the predictor variables but the structure otherwise stays the same: how should we do this?

>**Q.29.** Can you identify what effects are shown in the model summary?

::: {.callout-note icon=false collapse="true"}
- A.29. We can see that now the summary gives us estimates for the effects of:

- `FACTOR3_centered` -- this is the effect of reading strategy (`FACTOR3`) when `GENDER` level is `Female`
- `GENDERMale` -- this is the effect of `GENDER` when `FACTOR3_centered` is 0.

Note that the effect of `GENDER` is the estimated difference in outcome when `GENDER` level is `Female` (not shown, treated as the baseline) compared to when `GENDER` level is `Male`.

- `FACTOR3_centered*GENDER` -- this is the interaction, and reflects the change in the effect of `FACTOR3_centered` on outcome `mean.acc`, given different values of `GENDER`, here, when `GENDER` is `Male` instead of `Female`.
:::

- It helps to plot the effects estimates, given the model.

We can do this in a two-step sequence, similar to the sequence we trialled through weeks 17 and 18:

1. First fit the model
2. Second, use the model information to plot the predictions, given the effects

>**Q.30.** Can you identify how the effects of reading strategy (`FACTOR3`) and gender (`GENDER`) interact, given the model results summary and this plot?

::: {.callout-note icon=false collapse="true"}
- A.30. The plot, and the estimate for the interaction `FACTOR3_centered:GENDERMale`, together show that the coefficient for the effect of reading strategy (`FACTOR3_centered`) is slightly larger (the slope is slightly steeper) when `GENDER` is `Male` instead of `Female.

- Notice that, for this sample, the interaction is not significant, suggesting that there is no difference between genders in the relative effectiveness of reading strategy.
:::

## Online Q&A {#sec-wk19-lab-Q-and-A}

You will find, below, a link to the video recording of the Week 19 online Q&A after it has been completed.

<!-- ```{=html} -->
<!-- <iframe src="https://livelancsac-my.sharepoint.com/personal/shawej1_lancaster_ac_uk/_layouts/15/embed.aspx?UniqueId=85fcf27d-0a64-4d89-bf5b-250d3b42bc08&embed=%7B%22ust%22%3Atrue%2C%22hv%22%3A%22CopyEmbedCode%22%7D&referrer=StreamWebApp&referrerScenario=EmbedDialog.Create" width="640" height="360" frameborder="0" scrolling="no" allowfullscreen title="PSYC122 Online Q&A Week 11-14 & 16 -19-20250313_140242-Meeting Recording.mp4"></iframe> -->
<!-- ``` -->

