---
title: 3. DVs and IVs in RStudio
subtitle: Written by Tom Beesley (2025)
order: 4
---
## Week 3 Lecture

Watch Part 1

```{=html}
<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=6dcaa958-8ebb-4c7c-8aab-b36700ceaf8b&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=false&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="PSYC121 week 3 (A)" ></iframe>

```
Watch Part 2

```{=html}
<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=a2e69b40-b969-432a-a58f-b36700ced454&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=false&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="PSYC121 Week 3(B)" ></iframe>

```
Watch Lecture Part 3

```{=html}
<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=be0f83f6-501b-40a4-9365-b36700ceffab&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=false&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="PSYC121 week 3 (C)" ></iframe>

```
Watch Lecture Part 4

```{=html}
<iframe src="https://lancaster.cloud.panopto.eu/Panopto/Pages/Embed.aspx?id=e757e413-d17d-4085-8ef1-b36700cf1ef2&autoplay=false&offerviewer=true&showtitle=true&showbrand=true&captions=false&interactivity=all" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay" aria-label="Panopto Embedded Video Player" aria-description="PSYC121 week 3(D)" ></iframe>

```
Download the lecture slides [here](data/Week_3/week_3_lecture_slides_2025.pptx)

## Pre-lab work

Last week we asked you to

-   Use a script to run instructions in RStudio
-   Put data into RStudio form a data file and explore how to run descriptive stats and basic visualisations

This week - again, there's a learnr tutorial to follow and help prep for this week's activities. [You can find it here](https://ma-rconnect.lancs.ac.uk//Psych121_week_3/){target="_blank"}

## Lab activities

For the data concerning the weight of a cow data (week 2), we provided you with columns that reflected people's estimates, and you were able to generate descriptive statistics for those estimates (if you haven't done this, please go back and work through that part of the week 2 activity again). 

In order to find the estimates separately for the different groups of respondents (female and non-female), we needed to have a column for each gender category. Whilst that worked, it could get cumbersome over time always to work with data created like that.

*There is a better way...* In this first task we will look at how you can analyse the numerical weight estimates *as a function of* the categorical data in the identity column.

### Task 1 - More with the `penelope22` data

Step 1. Create a new folder for week 3, and set this as the working directory. This is covered in the learnr tutorial and we covered it in Week 2 as well. 

Step 2. Bring the `Week_3_2025.zip` file into R Studio server. Like last week, upload the zip file, and launch the R script. [You can get the file here](data/Week_3/Week3_2025.zip)

Step 3. This week, we again want to explore commands from the tidyverse library (toolkit) which can help us do more powerful things more elegantly. So let's get R to work again with the tidyverse library by running the code line

`library(tidyverse)`

Step 4. Explore help() commands. R can give you more information about how it works.

Step 5. Read the penelope22.csv data into R:

```{r eval=FALSE}
what_a_really_terrible_data_object_name <- read_csv(*MISSING*) 
# use your own data object name and specify the file you want to read in
```

Note that you will need to edit this line (and ensure you are in the correct working directory) for this to be successful. View the new data object using the View() command in the script or clicking on the object in the environment. 

This week we'll for the mean of the estimate data as a function of the different gender identities:

```{r eval=FALSE}
aggregate(x = *MISSING*$estimate, by = list(*MISSING*$identity), FUN = mean)
```

There's a lot going on in this code, but first, let's try replacing data object name where it says MISSING and then running the code. What do you get? Does this match what we did last week when we calculated the mean for the *female* and for the *other* (i.e., non-female) group?

Once you've run this, let's consider all the things that are going on in the code:

`aggregate` This is a command to call for descriptive statistics

`x=` This defines what column we are analyzing

`by=list` Now we tell R how to group the estimate data, and which column does that

`FUN=mean` Specifies which descriptive function is being asked for. Feel free to try functions for other descriptive statistics that you've used in previous weeks.

### using group_by() to calculate means for each group

Like most things in R, there are multiple ways to do the same thing. Here's another way to group scores by a (nominal/categorical) variable and calculate the values we need for each level of that variable. We explored this already in the learnr tutorial, and that should help you to complete the code below to get the weight estimates broken down by gender identity. You need to first define the data frame for the estimates data, and then specify the gender IV in the group_by command and the estimates DV in the summarise command:

```{r eval=FALSE}
*MISSING* %>% group_by(*MISSING*) %>% summarise(mean_estimate = mean(*MISSING*))
```

If you edit this code correctly and run it, you should get a result that is quite similar to that provided by the aggregate command (though the format of the output may be slightly different)

A note about: `%>%` 

This is called a "pipe operator", basically take the output from the left and feed it into the requests on the right. `Summarise` Provide summary statistics information for the specified variable as specified (whether mean, median etc)

### 3.1.3 The assignment operator

As well as learning about the pipe operator, we want to remind you /draw attention explicitly to another important element of the R command line syntax: the assignment operator. Using a command such as

`cows <- read_csv("penelope22.csv")`

looks for the csv datafile called 'penelope22', reads that data in, and assigns it to an object called 'cows'

We could create any object name we wanted (within broad limits of names that RStudio allows). The arrow symbol isn't just for reading in data files, we can perform a whole range of functions and assign them to an object.

#### Task 2 - New salary data

Using aggregate and summarise may not seem like much progress, because they are just replicating what we had already done with `mean()` is week 2. However (a) this emphasizes that there are often several ways to get at the same thing in R (b) now we know about doing calculations on grouped data, and about working with 2-dimensional data. These are big steps - we can now start to do much more efficient and informative things with our data.

Now, let's turn to the guesses made about median salary in the UK. We will read in some of the data that PSYC121 students provided in the file `wages2024.csv` (you will need to adapt the code we used above for the weight estimation file so that it will load in the wages data, and in what follows the assumption is your new variable name is called `wages`)

Let's take a peek at the dataset with:

`glimpse(wages)`

Glimpse pretty much does what you might think from the meaning of the word: it just gives us some quick statistics on the different column (handy because this is a much bigger dataset). We can see that we have 3 columns; uk_region (where someone lives, note 'other' probably equals Ireland, Europe, China, etc), family_position (age relationship with siblings), and salary (estimate).

We will be analysing how people assessed the median income in the UK. According to government statistics, the median income in 2023 was approximately £34,963  [see this link](https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/earningsandworkinghours/bulletins/annualsurveyofhoursandearnings/2023){target="_blank"}

Your task:

A)  Within the script, use the working "aggregate" commands from task 1 with the penelope weight data, can you find out the mean salary estimates as a function of where someone lives? That is, can you adapt that code you used earlier for this problem?

B)  Can you use the aggregate command to find out mean salary estimates as a function of the different family relationships? (if you are the youngest child maybe you have older siblings earning money that changes your evaluation?)

C)  Can you get a breakdown of salary estimates as a function of BOTH UK region AND family relationship together? You may well need some help with this, but have a guess at which bit of code would need to change to do it. 

D)  Can you use the `group by()` and `summarise()` command to display salary guesses as a function of where someone lives? Check this gives you the same answer.

## Task 3.2 - New phone use data

E)  Read in the dataset of phone screen time usage, `screentime2024.csv`, into a new object. For this task we'll focus on the `group_by()` and `summarise()` commands to further explore the data and consolidate your skills. Use copy and paste to adapt the existing script lines from the above tasks so that this time you calculate screen time usage as a function of the type of phone. 

F)  Use RStudio to figure out the (overall) mean screen time estimate and the standard deviation. Using these values, can you calculate by hand what screen time estimate value would reflect z scores of z=-1.5 and z = +2?

## Task 3.3 - Final challenge: visualisation

Can you find a way to visualise the screentime usage data that you have been working with above? The script provides two ways to consider doing this - basic boxplots (which we have looked at last week) and ggplot, which we have spent less time with but is an extremely powerful engine for creating plots. We've provided the start of the code in each case, leaving you to work out the specifics. 

# 4.  Now you're finished ...

Remember to add comments to your work in the script, and save the script file before you finish the session.


